### 2023-10-03 No Diamond for 3 layers.tex
No Diamond theorem
The first draft of No Diamond theorem

\documentclass[]{article}
\usepackage{amsmath, amsthm, amssymb, amsfonts}
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newcommand{\ReLU}{\mathrm{ReLU}}

\title{Proof part 1: 3 layers}
\date{}

\begin{document}

\maketitle

First we consider the case when there are only three layers: input layer, one hidden layer, and output layer. For convenience, we assume that the output only has one node. Of course, we assume all active functions are ReLU function.

We use $c_i$ to denote nodes in the input layer, use $b_j$ to denote nodes in the hidden layer and use $x$ to denote the output node. We use $W$ to denote the weights and $W_{bc}$ and $W_{xb}$ to denote the components. We use capital letter $B$ to denote bias, although it is not important here.

\begin{definition}
	1. We use $b$ to denote the node before ReLU and $\hat{b}$ to denote the node after ReLU: $\hat{b} = \ReLU(b)$.

	2. $\bar{f}$ is the upper bound approximation function of DeepPoly, and $\underline{f}$ is the lower bound approximation function.
\end{definition}

\begin{theorem}
	If there following conditions holds, then the bounds of $x$ obtained by DeepPoly are equal to the actual bounds:

	1. All DeepPoly triangle are right-angled triangle: for any ReLU node $b$ that upper bound $u(b)>0$ and lower bound $l(b)<0$, the approximation of lower bound is the zero function $\underline{f}(b)=0$.

	2. No Diamond: for any input node $c$, any output node $x$, any two paths $p_1,p_2$ from $c$ to $x$ along the direction of forward propagation. If $p_1$ is $c,b_i,x$, $p_2$ is $c,b_j,x$, then for $W_{b_jc}W_{xb_j}$ and $W_{b_ic}W_{xb_i}$, either one of them is $0$, or they have the same sign: both positive or negative.
\end{theorem}

\subsection*{One} First, we compute the bounds obtained by DeepPoly. Without loss of generality, we only consider the upper bound of $x$.   By the algorithm, the DeepPoly upper bound of $x$ is computed by following steps: \begin{align}
 \rightarrow&\sum_{W_{xb_i}<0}W_{xb_i}\underline{f_i}(b_i)+\sum_{W_{xb_i}>0}W_{xb_i}\bar{f_i}(b_i)+B_x\\
 \rightarrow&\sum_{W_{xb_i}<0}W_{xb_i}\underline{f_i}(\sum_{c_j}W_{b_ic_j}c_j+B_i)+\sum_{W_{xb_i}>0}W_{xb_i}\bar{f_i}(\sum_{c_j}W_{b_ic_j}c_j+B_i)+B_x\\
 =& C+\sum_{W_{xb_i}<0}W_{xb_i}\cdot\underline{k_i}\cdot(\sum_{c_j}W_{b_ic_j}c_j+B_i)+\sum_{W_{xb_i}>0}W_{xb_i}\cdot\bar{k_i}\cdot(\sum_{c_j}W_{b_ic_j}c_j+B_i)\\
 =& C+\sum_{c_j}c_j\cdot\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i\\
 \rightarrow& C+\sum_{\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i<0}\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i\cdot l(c_j)\\
 &+\sum_{\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i>0}\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i\cdot u(c_j)
\end{align} Here, $\bar{k_i}$ is the coefficient of $\bar{f_i}$, which is a linear function, and $\underline{k_i}$ is the coefficient of $\underline{f_i}$, which is also a linear function. $C$ is the sum of all constants that occurs in above computations. $K_i$ is the coefficient $\bar{k_i}$ or $\underline{k_i}$ depending on $i$. Notice that by the definition of DeepPoly, all $K_i$ are non negative.

Among these formulas, (1) to (4) are abstract formula with formal variables $b_i,c_j$. The result of (5)(6) is the upper bound of DeepPoly for node $x$.

For for each fixed $c_j$, because we assumed no Diamond, so all (for different $b_i$) $W_{xb_i}W_{b_ic_j}$ must have the same sign, or be $0$, and so is $\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i$. So we assign each node $c_j$ a sign of $+,-,0$,  if at least one of $W_{xb_i}W_{b_ic_j}$ is positive, at least one of $W_{xb_i}W_{b_ic_j}$ is negative, or all $W_{xb_i}W_{b_ic_j}$ are 0. Then we use $s(c_j)$ be $u(c_j)$ or $l(c_j)$ or $0$ if $c_j$ has sign of $+$ or $-$ or $0$.

Then (5)(6) can be simplified as \begin{align}
	=C+\sum_{c_j}s(c_j)\cdot\sum_{b_i}W_{xb_i}W_{b_ic_j}\cdot K_i
\end{align} This is the DeepPoly upper bound of $x$.

\subsection*{Two} Second, we show that when each $c_j$ of $+,-$ sign gets $s(c_j)$ value, the actual value of $x$ is the same as its DeepPoly upper bound, hence the DeepPoly upper bound is the actual upper bound.

Using the chain from (4) to (2), (7) is equal to: \begin{align*}
&\sum_{W_{xb_i}<0}W_{xb_i}\underline{f_i}(\sum_{c_j}W_{b_ic_j}s(c_j)+B_i)\\
+&\sum_{W_{xb_i}>0}W_{xb_i}\bar{f_i}(\sum_{c_j}W_{b_ic_j}s(c_j)+B_i)+B_x
\end{align*}

We only need to show when all $c_j$ of $+,-$ sign get $s(c_j)$ value, each $\bar{f_i}(\cdots)$ in above formula (or $\underline{f_i}$) is equal to the actual value of $\hat{b}_i$.

\subsubsection*{positive} We fix a node $b_i$ that $W_{xb_i}>0$. If so, every $W_{b_ic_j}$ has the same sign as $W_{x_bi}W_{b_ic_j}$, and hence has the same sign as $c_j$ or is $0$ (and if $c_j$ has sign $0$, then $W_{b_ic_j}=0$, too). Consider all nonzero $W_{b_ic_j}$: because \begin{align}
	b_i = \sum_{c_j} W_{b_ic_j}c_j+B_i,
\end{align}  when all $c_j$ of $+,-$ sign get $s(c_j)$ value, $b_i$ also gets its upper bound. Because $b_i$ is in the second layer, this upper bound is equal to the DeepPoly upper bound of $b_i$.

Moreover, by the algorithm of DeepPoly, for any case of $\bar{f_i}$, if $b_i$ get its DeepPoly upper bound, then $\bar{f_i}(b_i)=\ReLU(b_i)$.

Put all together, when all $c_j$ of sign $+,-$ gets values $s(c_j)$, $b_i=\sum_{c_j}W_{b_ic_j}s(c_j)+B_i$ gets its actual upper bounds and DeepPoly upper bounds, and then  $\bar{f_i}(b_i)=\ReLU(b_i)$, the actual value of $\hat{b}_i$. This is what we want to show.

\subsubsection*{negative}Now we consider a node $b_i$ that $W_{xb_i}<0$. All are similar. Now, each $W_{b_ic_j}$ has the opposite sign of $c_j$. So when all $c_j$ of $+,-$ sign get $s(c_j)$ value, $b_i$ also gets its lower bound and DeepPoly lower bound. By the assumption, when $b_i$ gets its DeepPoly lower bound, $\underline{f_i}(b_i)=\ReLU(b_i)$.

151+124+103+48+3(estimate)=428

0 : 1, 6.942969230753979, 933.6970219612122

open 30 : 59 : -1, 1.3304800987255705, 213.07373809814453, DiamondMin of Min: -2.4831832946871297

open 0:   59 : -1, 2.2208665044849285, 110.11069464683533, DiamondMin of Min: -3.2142924202998118

open 20:  59 : -1, 1.5128416712774397, 169.37895846366882, DiamondMin of Min: -2.5436423082465973

open 40: 59 : -1, 1.1588057492305743, 167.29228901863098, DiamondMin of Min: -2.322909620171307

open 50: 59 : -1, 1.1332793237397172, 216.03394603729248, DiamondMin of Min: -2.3194345600969135

open 50: 59 : -1, 1.136410977665245, 250.3934416770935, DiamondMin of Min: -2.3194345600969135

open 50 (more data): 59 : -1, 1.1033637386628505, 321.2751393318176, DiamondMin of Min: -2.279701100000497

open 10: 59 : -1, 1.7965610184250298, 144.45073246955872, DiamondMin of Min: -2.8121700263290466

%random open 10: 59 : -1, 2.0868988971016167, 133.93379545211792, DiamondMin of Min: -3.070953068366777

%random open 20: 59 : -1, 1.9800122543451073, 130.6924500465393, DiamondMin of Min: -3.0701026929909143
%
%random open 20: 59 : -1, 1.97995359889014, 135.06353521347046, DiamondMin of Min: -2.9642806539605853

random open 21: 59 : -1, 1.8202823109335369, 158.56950044631958, DiamondMin of Min: -2.926562550850029

random open 21: 59 : -1, 1.8113843758166142, 128.9297330379486, DiamondMin of Min: -2.986175395636058

random open 31: 59 : -1, 1.617607152194705, 142.7976849079132, DiamondMin of Min: -2.7544434938921665

random open 31: 59 : -1, 1.6338272022581422, 147.04282569885254, DiamondMin of Min: -2.734869999570412

random open 41: 59 : -1, 1.4376834177337172, 141.38413166999817, DiamondMin of Min: -2.591134103101991

random open 51: 59 : -1, 1.2285876079490041, 218.22643494606018, DiamondMin of Min: -2.461825229746404

random open 61: 59 : -1, 1.0442270697364395, 315.7167875766754, DiamondMin of Min: -2.219210859797969

random open 61: 59 : -1, 1.0458496294859592, 375.4530429840088, DiamondMin of Min: -2.219210859797969

\end{document}