### 2024-01-24 Paper/proofs.tex
proof

\section*{Appendix}
As a first step to prove Theorems \ref{th1},\ref{th2}, 
we will relax the compensating pairs notion to allows pairs of paths having nodes in common.
Such a pair is called a {\em weak compensating pair}.
We will then in a second step prove Theorems \ref{th1},\ref{th2} for the full definition of compensating pairs of paths.

Let $\cal B$ be a neighbourhood of an input.
For convenience,
when the neighbourhood $\cal B$ is clear, 
we will write $\max(n)$ to denote $\max_{\vx \in \cal B}(\Val_{\vx}(n))$, and $\min(n)$ to denote $\min_{\vx \in \cal B}(\Val_{\vx}(n))$.
Notice that if $\Val_{\vx}(b)$ is maximal (resp. minimal), 
then $\Val_{\boldsymbol{x}}(\hat{b})$ also gets maximal (resp. minimal).

We fix a target node $z$.
We now define the partial sign function $S^z$, denoted $S$ when $z$ is clear from the context:

\begin{definition}
	\label{sign_of_nodes}
	We define the partial sign function $S$ on node $n$: 	
	\begin{enumerate}
		\item $S(n)=0$ if all path from $n$ to $z$ have weight $0$; 
		\item $S(n)=1$ if all path from $n$ to $z$ have non-negative weight, and at least one path has a positive weight; 
		\item $S(n)=-1$ if all path from $n$ to $z$ have non-positive weight, and at least one path has a negative weight. 
	\end{enumerate}
\end{definition}

In general, $S$ may not be defined on every node 
(e.g. if there is a negative and a positive path from $a$ to $c$). 
However, if there is no compensating pair of paths, 
$S$ is defined on all nodes: any node fulfills one of the above cases (1),(2),(3).

For instance for any 3 layer DNN: for an input node $a_i$:
\begin{enumerate}
	\item  $S(a_i)=0$ if 
	for all $b_j, W_{a_i b_j}\cdot W_{b_j c} = 0$

	\item  $S(a_i)=1$ if for all $b_j, W_{a_i b_j}\cdot W_{b_j c} \geq 0$ and there exists 
	$j, W_{a_i b_j}\cdot W_{b_j c} > 0$

	\item $S(a_i)=-1$ if for all $b_j\ W_{a_i b_j}\cdot W_{b_j c} \leq 0$ and there exists 
	$j, W_{a_i b_j}\cdot W_{b_j c} < 0$ 
\end{enumerate}

For $b_j$ in the hidden layer, we have $S(b_j)=1,-1,0$ if $W_{b_j c}$ is positive, negative, or 0 respectively. Finally, for the output node $z$, we define $S(z)=1$.

\section{Proof of Theorem \ref{th1}}

Assume that there is a no (weakly) compensating pair of paths.

\begin{lemma}
	\label{lemma1}
	Let $L,L'$ be consecutive layers of a DNN without compensation, 
	$m\in L$ and $n\in L'$. If $W_{m n} \neq 0$ and $S(n) \neq 0$, then 
	$S(m)=S(n)\mathrm{Sign}(W_{n m})$.
\end{lemma}

\begin{proof}
	If $S(n) \neq 0$, then there is a path $\pi$ from $n$ to the output node $z$ with a nonzero weight of the same sign as $S(n)$. 
	So there is a non zero path from $m$ to $z$: $(m n) \pi$, whose sign is 
	$S(n)\mathrm{Sign}(W_{n m})$. As there is no compensation, $S(m)=S(n)\mathrm{Sign}(W_{n m})$.
	\hfill $\square$
\end{proof}

%For a node $n$, we use $n_s$ to denote $S(n)\cdot n$. 
%Notice that for $S(n)=1$, $n_s$ gets maximal value whenever $n$ gets maximal value; 
%while for $S(n)=-1$, $n_s$ gets maximal value whenever $n$ gets minimal value (and vice versa). For $S(n)=0$, $n_s=0$ and thus always reach his minimal and maximal.

Using the concept of sign, now we introduce the key lemma for Theorem \ref{th1}.

\begin{definition}
We define the following two input vectors $\vx^{*}, \vx^{\sharp}$ in $\cal B$: 
	\begin{itemize}
		\item $\vx^{*}$ is the input vector defined in the following way:
		\begin {itemize}
		 \item $x^*_{a_i}=\max(a_i)$ if $S(a_i)\in \{0,1\}$, and
          \item $x^*_{a_i}=\min(a_i)$ otherwise, that is when $S(a_i)=-1$.
	\end{itemize}

		\item $\vx^{\sharp}$ is the input vector defined in the following way:
		\begin{itemize}
			\item $x^{\sharp}_{a_i}=\min(a_i)$ if $S(a_i)\in \{0,1\}$, and
			\item $x^{\sharp}_{a_i}=\max(a_i)$ otherwise, that is when $S(a_i)=-1$.
		\end{itemize}
	\end{itemize}

\end{definition}

\begin{lemma}
	\label{lem:key1}
	\label{lemma2}
For all $n$ with $S(n)=1$, we have $\val_{\vx^*}(n) = \max(n)$
and $\val_{\vx^{\sharp}}(n) = \min(n)$.
For all $n$ with $S(n)=-1$, we have $\val_{\vx^*}(n) = \min(n)$
and $\val_{\vx^{\sharp}}(n) = \max(n)$. Further, for $L,L'$ two consecutive layers and $n$ on layer $L'$:
$$\max(n)=\sum_{m \in L, W_{mn}>0}W_{mn} \max(\hat{m}) + \sum_{m \in L, W_{mn}<0}W_{mn} \min(\hat{m}) + bias_n$$
$$\min(n)=\sum_{m \in L, W_{mn}>0}W_{mn} \min(\hat{m}) + \sum_{m \in L, W_{mn}<0}W_{mn} \max(\hat{m}) + bias_n$$

\end{lemma}

\begin{proof}
	We prove this lemma by induction on layers. For input layer, this is true by definition. Suppose this lemma is true up to layer $L_i$, we need to show this for layer $L_{i+1}$. Let $n$ be a node in layer $L_{i+1}$, then by definition, we have the following equation: \begin{align*}
		\val_{\vx}(n) = \sum_{m\in L}W_{m n}\val_{\vx}(\hat{m})) + bias_n.
	\end{align*}

	By induction hypothesis, when input vector is $\vx^{*}$, for every $m$, if $S(m)=1$, then $\Val_{\vx^*}(\hat{m})=\max(\hat{m})$ and if $S(\hat{m})=-1$, then $\Val_{\vx^*}(\hat{m})=\min(\hat{m})$. Assume that $S(n)=1$. By applying Lemma \ref{lemma1} and the induction hypothesis:
	\begin{align*}
		\val_{\vx^*}(n) = \sum_{m \in L_i, W_{mn}>0}W_{mn} \max(\hat{m}) + \sum_{m \in L_i, W_{mn}<0}W_{mn} \min(\hat{m}) + bias_n\\
	\end{align*}

	Further, we have the following inequality:
	\begin{align*}
		\max(n)\leq & \sum_{m \in L_i, W_{mn }>0}W_{mn} \max(\hat{m}) + \sum_{m \in L_i, W_{mn}<0}W_{mn} \min(\hat{m}) + bias_n\\
		\leq & \val_{\vx^*}(n)
	\end{align*}

	%If $S(m)=1$, then by Lemma \ref{lemma1}, $W_{mn}>0$; if $S(m)=-1$, then $W_{mn}<0$; if $S(m)=0$, then $W_{mn}=0$.

	By definition of $\max$, we have $\val_{\vx^*}(n) \leq \max(n)$.
	Combining the 2 formulas, we obtain $\max(n)=\Val_{\vx^*}(n)$.

	We prove similarly that $\min(n)=\Val_{\vx^\sharp}(n)$,
	and the case $S(n)=-1$. This concludes the induction. \qed
\end{proof}

\iffalse
	As direct corollary of Lemma \ref{lem:key1}, we obtain:

	\begin{corollary}
		\label{cor1}
		$$\max(c)=\sum_{b \in L, W_{bc}>0}W_{bc} \max(\hat{b}) + \sum_{b \in L, W_{bc}<0}W_{bc} \min(\hat{b}) + bias_c$$
		$$\min(c)=\sum_{b \in L, W_{bc}>0}W_{bc} \min(\hat{b}) + \sum_{b \in L, W_{bc}<0}W_{bc} \max(\hat{b}) + bias_c$$
	\end{corollary}

\fi

	%Notice that for all $x$, for either $S(c)=1$ or $S(c)=-1$, these two bounds are found in Lemma \ref{lemma2},
	%from configuration $\boldsymbol{x}^*$ and $\boldsymbol{x}^\sharp$.

	\subsubsection*{Bounds generated by Value Abstraction}

	\smallskip

	Let $[\alpha_n,\beta_n]$ be the bounds generated by the Box abstraction.
	We show that these bounds are exact when there is no (weak) compensating pairs of paths, i.e. for all node $n$ of the DNN, we have $\alpha_n=\min(n)$ and $\beta_n=\max(n)$.

\begin{proof}
	 We show that inductively. The initialization is obvious as Box Abstraction is always exact for the first layer.

	Let $n$ be node of the DNN. We set the target $z=n$ and consider the sign function $S^z=S^n=S$ corresponding to that node. By definition, $S(n)=1$.

	The Box abstraction computes its upper bound using:
	$$\beta_n= \sum_{W_{mn}>0} W_{mn} \beta_{\hat{m}} + \sum_{W_{mn}<0} W_{mn} \alpha_{\hat{m}} + bias_n$$

	By induction hypothesis, we have 
		$\beta_{\hat{m}}=\max(\hat{m})$ and
		$\alpha_{\hat{m}}=\min(\hat{m})$, thus 
		applying the second part of Lemma \ref{lem:key1}, we obtain
		$\beta_n=\max(n)$.

		Similarly, we obtain $\alpha_n=\min(n)$.

		Consider now $\alpha_{\hat{n}},\beta_{\hat{n}}$.
		If $\alpha \leq 0$ then $\alpha_{\hat{n}}=0=\max(\hat{n})$.
		Otherwise, $\alpha_{\hat{n}}=\alpha_n=\max(\hat{n})$.
		In both case it is exact.

		Similarly, if $\beta_n=\max(n)<0$, 
		then $\beta_{\hat{n}}=\max(\hat{n})=0$, 
		and otherwise 
		$\beta_{\hat{n}}=\max(\hat{n})=\beta_n=\max(n)$.

		Hence we have equality pre and post activation function. \qed

			\end{proof}

This shows the statement of Theorem 1 with the weaker statement that there is no weak compensating pairs of paths.

	\subsubsection*{From weak to normal}

	\begin{lemma}
		If a DNN has no compensating pair, then it also does not have weak compensating pair from an input node to the output node.
	\end{lemma}

	\begin{proof}
		By contradiction, if there is a weak compensating pair of two paths, $\rho_1,\rho_2$, 
		then consider the common nodes by increasing layer $n_1 \cdots n_k$ in $\rho_1,\rho_2$ (including the initial and final nodes).
		Assume that there is no compensating pair.
		It means that all the segment of $\rho_1,\rho_2$ from $n_i$ to $n_{i+1}$ have the same sign (they have no nodes in common but the initial and final nodes).
		Hence $\rho_1$ and $\rho_2$ have the same sign, a contradiction with $\rho_1,\rho_2$ being a weakly compensating pair.  \qed
	\end{proof}

	This finishes the proof of Theorem \ref{th1}.

	\newpage

\section{Proof of Theorem \ref{th2}}

			In this section, we will prove theorem \ref{th2}. We first prove the following weaker theorem:

			\begin{theorem} \label{thm:2}

				Suppose $\mathcal{D}$ is a DNN with $\ReLU$ as the unique activation functions and with a unique output node $z$. 
				Let $X$ be the set of nodes $n$ such that there is a weak compensating path 
				passing by $n$ (except its source and target nodes).
				Then $\MILP_X$ computes the exact bounds $[\alpha,\beta]$ for $z$,
				i.e. $\alpha=\min(z)$ and $\beta=\max(z)$.
			\end{theorem}

			\subsection{Preparation}

			\begin{definition}
				We define a partition of the set $A$ of input nodes as $A_{zero} \sqcup A_{pure}\sqcup A_{open}$, with:  
				\begin{itemize}
					\item $A_{zero}= \{a \mid \forall \text{ path $\rho$ from $a$ to } z, weight(\pi)=0\}$.

					\item $a_k \in A_{pure}$  if $a \notin A_{zero}$ and either
					\begin{enumerate}
						\item every path from $a_k$ to $z$ has weight $\geq 0$, or
						\item every path from $a_k$ to $z$ has weight $\leq 0$.

					\end{enumerate}
					That is, $a_k$ is not a source of a weak compensating pair. We denote 
					$A_{pure}$ by $I$ for short.
					\item $a_k \in A_{open}$ iff $a_k \notin A_{pure} \sqcup A_{zero}$, that is there exists two paths $\rho,\rho'$ from $a_k$, one with positive and one with negative weight. 	We denote $A_{open}$ by $J$ for short.
				\end{itemize}
			\end{definition}

			\begin{lemma} \label{lem:open_node_2}
				A node $b$ in a hidden layer will not be on a weak compensating path if and only if one of the following happens:
				\begin{enumerate}
					\item Any path from $b$ to the output node $z$ has a weight $0$, and we use $weight({bz})=0$ to denote this case, or
					\item For every input node $a_j\in J$, every path from $a_j$ to $b$ has a weight of $0$. We use $weight({a_jb})=0$ to denote this case.
				\end{enumerate}

			\end{lemma}

			We denote $B_{zero}$ the set of nodes $b$ that case 1 holds, and denote $B_{pure}$ the set of nodes $b$ such that case 2 holds but case 1 does not hold.
			We denote $B_{open}$ the set of all nodes $b$ in a weak compensating path (remaining cases).

			\begin{proof}
				First, we show that if one of case 1, 2 happens, then $b$ will not be in a weak compensating pair. If 1), it is obvious. For 2, we reason by contradiction: assume there is a pair of weak compensating paths 	$(\pi,\pi')$ starting with $a$, with $b$ in $\pi$ and weight$(\pi) > 0$. It means that $a \in J$. This is a contradiction as 2) $weight({a_j,b})=0$ implies weight$(\pi)=0$.

				Second, we show that if neither 1 nor 2 hold, then $b$ will be in a weak compensating path.
				Because 2) does not hold, there exists $a_j \in J$ with $weight({a_j,b}) \neq 0$, say $>0$.
				Because 1) does not hold, $weight({b,z}) \neq 0$, say $>0$.
				Now, by definition of $J$, there is a pair of weak compensating paths $\pi,\pi'$ 
				from $a_j$ to $z$, say with $\pi'$ with weight $<0$.
				Then two paths $(a_j,b,z)$ and $\pi'$ also form a weak compensating pair. \qed
			\end{proof}

			\begin{lemma}\label{lem:subnetwork}
				$B_{pure}$ and $I$ form a sub-network, that is, 
				for $n \in B_{pure}$, for every path $\rho$ from $m$ to $n$
				either $m \in B_{pure}\cup I$ or $weight(\rho)=0$.
			\end{lemma}

			We call $D_I$ the subnetwork  made of nodes in $B_{pure}\cup I$.
			In $D_I$, there is no weakly compensating pair.

			\begin{proof}
				This is simply because for a node $c\in B_{pure}$ in a hidden layer or the output layer, for a node $b$ in one layer before, if $b\notin B_{pure}\cup I$, then:

				1. if $b\in J$, by definition, $W_{bc}=0$.

				2. if $b\in B_{open}$, then $W_{bc}=0$; otherwise there exists $<a_j,b,c>$ a path from $a_j\in J$ to $c$ with a nonzero value, a contradiction.

				3. if $b\in B_{zero}$, then $W_{bc}=0$; otherwise there exists $<b,c,z>$ a path from $b$ to $z$ the output node with a nonzero value, a contradiction. \qed
			\end{proof}

			Recall the sign function of nodes from Definition \ref{sign_of_nodes}. Because weak compensating paths are allowed, the sign function is no longer defined on all nodes. $S$ is defined on all nodes $a_i \in I$. 
			Further, we have:

			\begin{lemma}\label{lem:sign}
				For a node $n$, if there exists at least one path from a node $a_i\in I$ to $n$ with nonzero weight, then $S$ is defined on $n$.
			\end{lemma}

			\begin{proof}
				By contradiction, suppose that there exists two paths $\rho_1,\rho_2$ from $n$ to $z$ with (strictly) different signs. The pair $\langle a_i,n\rangle \cdot \rho_1$ and $\langle a_i,n\rangle \cdot \rho_2$ is compensating, a contradiction with $a_i\in I$. \qed
			\end{proof}

			Hence for any node $m$ such that $S$ is undefined on $m$, value of $m$ can only depends upon input nodes in $J$.

			\iffalse
			\begin{lemma}
				If $n$ is in the layer after $m$, then $S(n), S(m)$ are both defined
				with $S(n) \neq 0$ and $W_{mn}\neq 0$, then $sign(W_{mn})S(n)=S(m)$.
			\end{lemma}

			\begin{proof}
				This is proved in the first section of appendix. \qed
			\end{proof}

			\fi

			\begin{lemma}\label{lem:subnetwork2}
				The input set $J \cup B_{open}$ forms a sub-network, denoted $D_J$. 
				$S$ is undefined on all nodes of $D_J$.
			\end{lemma}

			\begin{proof}
				This is because if there exists nodes $b,c$ such that $W_{bc}\neq 0$ and $c$ is in above set but $b$ is not, then $b$ has a path with nonzero weight from a node $a_i$ not in $J$ but this means so does $c$, a contradiction. Therefore any $b$ such that $W_{bc}\neq 0$ is also in above set. \qed
			\end{proof}

			By definition, for any $b\in D_J$, either $b\in B_{open}$, or $b$ is a constant node: any path from any input node must have zero weight to it.

			We denote $\boldsymbol{x}_S$ for any subset $S$ of input nodes to refer the input vector $\langle x_{a_k}\rangle_{a_k\in S}$. We also use $\boldsymbol{x}_I\oplus \boldsymbol{x}_J = \boldsymbol{x}$ and $\Val(z)=\Val_{\boldsymbol{x}}(z)=\Val_{\boldsymbol{x}_I,\boldsymbol{x}_J} (z)$. For a node $b$ that $S$ is not defined on it, $b$ only depends on $J$; so, if $\boldsymbol{x}_J$ is fixed, then the value of $b$ will be fixed.

			We define $\boldsymbol{x}_I^*$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=-1$). Similarly, we define $\boldsymbol{x}_I^\sharp$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=-1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=1$).

			\begin{lemma} \label{lem:reach_max_2}

				Assume $b$ is the output node or a node in hidden layers that $S(b)=1$, then   for any valuation $\boldsymbol{x}^0_J$, 
				we have: $$\max_{\boldsymbol{x}_I} (\Val_{\boldsymbol{x}_I,\boldsymbol{x}_J=\boldsymbol{x}^0_J}(b)) =  \Val_{\boldsymbol{x}_I^*,\boldsymbol{x}^0_J}(b).$$

				If $S(b)=-1$, then the similar results hold, replacing $\boldsymbol{x}^*_I$ by $\boldsymbol{x}^\sharp_I$, or replacing $\max$ by $\min$. 
			\end{lemma}

			\begin{proof}

				For any fixed input vector $\boldsymbol{x}^0_J$, we can regard $\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z)$ as a DNN $D'$ with input nodes $a_i\in I$, with all $\boldsymbol{x}^0_J$ %and their propagation in hidden layers 
				as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Lemma \ref{lem:key1} to get that $\Val_{x_I,\boldsymbol{x}^0_J}(z)$ reaches its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$, and we are done. \qed
			\end{proof}

			\subsection{MILP abstraction}

			The standard MILP abstraction use the abstraction from \cite{MILP}, 
			where each variable $\alpha_b$ for the ReLU from $b$ to $\hat{b}$ is:
			\begin{itemize}
				\item linear  (i.e. $\alpha_b \in [0,1]$) for $b \notin B_{open}$
				\item binary/integer (i.e. $\alpha_b \in \{0,1\}$) for $b \in B_{open}$
			\end{itemize}

			Now consider the MILP framework for all hidden and output nodes layer by layer (to compute their bounds, and use the bounds of previous layers to build the MILP models of current layer), with ReLUs being either linear or binary depending on whether $b \in B_{open}$ or not. We call this standard framework.

			For a node $b$ in a hidden layer or the output layer, we denote by $\UB(b)$ the upper bound on $b$ and $\LB(b)$ the lower bound on $b$ considering the MILP constraints above.

			\begin{lemma}
				In the standard $MILP$ framework, both $D_I$ and $D_J$ are accurate in the sense that for any $b$ in $D_I$ or $D_J$, $\UB(b)=\max(b)$ and $\LB(b)=\min(b)$.
			\end{lemma}

			\begin{proof}
				For $D_I$, it is because of no compensating as proved in the first section. For $D_J$, it is because all nodes are either opened or constants. \qed
			\end{proof}

			%\begin{definition} Assume $\boldsymbol{x}^0_I\oplus \boldsymbol{x}^0_J$ is a input vector, and $b$ is node in hidden layers or output layer.
			%
			%	1. For a given vector $\boldsymbol{x}$ , $\B_{\boldsymbol{x}^0_I, \boldsymbol{x}^0_J, \boldsymbol{x}}(b)$ is the value of $b$ in the MILP model for values $\boldsymbol{x}^0_I, \boldsymbol{x}^0_J, \boldsymbol{x}$.
			%\end{definition}

			For input vectors $\boldsymbol{x}^0_I,\boldsymbol{x}^0_J$, let $\UB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)$ be the upper bound in the above fixed formulation and framework while the input $\boldsymbol{x}^0_I,\boldsymbol{x}^0_J$ is fixed: that is, besides the constraints in above MILP models, we add new constraints $\boldsymbol{x}_I,\boldsymbol{x}_J=\boldsymbol{x}^0_I,\boldsymbol{x}^0_J$. We  define $\LB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)$ similarly for $\LB$ lower bound. For them, we have the following lemma:

			\begin{lemma} In an MILP formulation, for a node $c$ in a hidden layer or output layer:

				1. $\UB(b)=\max_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}\UB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(c)$.

				2. $\LB(b)=\min_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}\LB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(c)$. 
			\end{lemma}

			\begin{proof}
				This is by the definition of MILP formulation and $\UB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}, \LB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}$. \qed
			\end{proof}

			We want to show that $\mathrm{UB}(z) = \max (z)$ and $\mathrm{LB}(z) = \min (z)$ for the unique output nodes. In fact, we will prove this for all $c\notin B_{zero}$ . As the MILP abstraction is a sound over-approximation, 
			it suffices to show that $\mathrm{UB}(c)\leq \max (c)$ and $\mathrm{LB}(c) \geq \min (c)$. We will prove this by induction on layers from the first hidden layer to the output layer. More specifically, we will prove the following lemmas:

			\begin{lemma} \label{lem:main}
				For any node $c$ in a hidden layer or the output layer, if $c\notin B_{zero}$, then:		\begin{enumerate}
					\item for any fixed $\boldsymbol{x}^0_J$, when  $\boldsymbol{x}_I=\boldsymbol{x}^*_I$, the value for $c$ in the MILP model is a fixed number, and naturally $\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)=\LB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)$. The same is true for $\boldsymbol{x}^\sharp_I$.

					\item for any fixed $\boldsymbol{x}^0_J$, if $S(c)=1$, then $$\max_{\boldsymbol{x}_I} \UB_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)=\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)= \Val_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c) = \max_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c).$$ The similar for $S(c)=-1$, $\langle \LB,\min\rangle$, $x^\sharp_I$: we can change two of them at the same time.

					\item Therefore, $\UB(c)=\max(c)$ and $\LB(c)=\min(c)$.
				\end{enumerate}

			\end{lemma}

			\begin{proof}
				We prove this lemma by induction on layers. For first hidden layer $L_1$, it is obvious. Suppose we have proved part 1,2,3 up to layer $L_i$, then for layer $L_{i+1}$, suppose $c$ is a node in this layer.

				(1)	If $S(c)$ is undefined, then by previous discussion, $c$ is in a sub-network from $J$ with all nodes opened; and if $\boldsymbol{x}_J=\boldsymbol{x}_J^0$ is fixed, then in the MILP model, the value for $c$ is also fixed.

				If $S(c)$ is defined, then consider all $b$ such that $W_{bc}\neq 0$. First, $b$ cannot in $B_{zero}$; so by induction hypothesis, the value of $b$ is fixed under $\boldsymbol{x}^*_I,\boldsymbol{x}^0_J$ in MILP model.

				If $b\in B_{open}$, the term is $W_{bc}\ReLU(b)$. We know then by induction hypothesis, the value for $b$ in MILP is a fixed value. Then the term $W_{bc}\ReLU(b)$ is also a fixed number.

				If $b\in B_{pure}$, the term is $W_{bc}\mathrm{AppB}(b)$ ($\mathrm{AppB}$ is the LP approximation domain of $\ReLU$) and either $S(b)$ is undefined and $b$ is a constant in the network and so is in the MILP model (trivial case), or $S(b)$ is defined.

				Then by induction hypothesis for all 1, 2 and 3 parts,  the value of $b$ in the model is fixed (part 1), and it is equal to $\max_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)$ or $\min_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)$ (by part 2, depending on $S(b)$). Since $b$ only depends on $I$, this value also equal to $\max_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I}(c)$ or $\min_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I}(c)$, and by induction hypothesis on part 3, equal to $\UB(b)$ or $\LB(b)$. For either of them, $\mathrm{AppB}(b)$ will be fixed value. Hence, in this case, the term $W_{bc}\mathrm{AppB}(b)$ is also a fixed number.

				Therefore, in the MILP model, $c$ is a sum of terms that are all fixed numbers. Hence $c$ is a fixed number.

				(2) If $S(c)=1$, then we have the following:	\begin{align*}
					\mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^0_I}(c) \leq constant + \sum_{b\in B_{open}\cap l_i} W_{bc}\ReLU(\B_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}} W_{bc} \mathrm{AppB}(\B_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)),
				\end{align*}where $\B$ means one of $\UB$ or $\LB$ depending on the sign of $W_{bc}$: if $W_{bc}$ is positive, then $\UB$, otherwise $\LB$. Notice that all $b$ displayed are that $S(b)$ defined, because if $S(b)$ undefined, then for fixed $\vx_J$, value of $b$ is fixed and will be put into the constant.

				Now by induction hypothesis part 2, if $S(b)=1$, then $\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)=\max_{\vx_I}\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)$; and if $S(b)=-1$, $\LB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)=\min_{\vx_I}\LB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)$. Moreover, since $S(c)=1$, $S(b)$ and $W_{bc}$ have the same sign. And both $\ReLU$ and $\mathrm{AppB}$ are non-decreasing functions. Combining all these facts, we have:\begin{align*}
					\mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^0_I}(c)\leq 
					&constant + \sum_{b\in B_{open}\cap l_i} W_{bc}\ReLU(\B_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}} W_{bc} \mathrm{AppB}(\B_{\boldsymbol{x}^*_I}(b))\\
					= & \mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^*_I}(c). 
				\end{align*}  This is the first equal.

				The second is simply because $\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)\geq \Val_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)\geq\LB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)$.	The third equal is proved in last subsection, Lemma \ref{lem:reach_max_2}. The case $S(c)=-1$ is similar. This is for part 2.

				(3) Once we have part 2, we will have that for any fixed $\boldsymbol{x}^0_J$, $$\max_{\boldsymbol{x}_I} \UB_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)= \max_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I,\boldsymbol{x}_J}(c).$$ Therefore, $\max_{\boldsymbol{x}_I,\boldsymbol{x}_J} \UB_{\boldsymbol{x}_I,\boldsymbol{x}_J}(c)= \max_{\boldsymbol{x}_I,\boldsymbol{x}_J} \Val_{\boldsymbol{x}_I,\boldsymbol{x}_J}(c)$ and this is to say $\UB(c)=\max(c)$. This is what we want to show, and the same result holds for lower bound and minimal value. This is for part 3. \qed

			\end{proof}

			From Lemma \ref{lem:main}, we already have that for the unique output node $z$, $$\UB(z)=\max(z).$$ This finish the proof of Theorem \ref{thm:2}.

			\subsubsection*{From weak to compensating}

			Now, we show that Theorem \ref{thm:2} can prove Theorem \ref{th2}. Suppose $a$ is a node in a weak compensating path $P$, and $P,P'$ form this weak compensating pair.

			Cut $P,P'$ into segments $P_0,P_1,\cdots$, and $P'_0,P'_1,\cdots$ at each common node; hence each $P_i,P_i'$ either have not common nodes (except the start and end nodes), or are the same. Now, for node $a$, there are three cases:
			\begin{itemize}
				\item $a$ is a common nodes.

				\item  $a$ is a in path $P_i$ such that weights of $P_i,P_i'$ have the same sign.

				\item  $a$ is a in path $P_i$ such that weights of $P_i,P_i'$ have the different sign.
			\end{itemize}

			In the last case, $a$ is in a compensating path (the difference is that this path is not from the input to output), and will be opened. So we focus on the first and second case.

			For the second case, if we replace $P'$ with $P''=P'_0,\cdots,P'_{i-1},P_{i},P'_{i+1},\cdots$, then $P,P''$ is a weak compensating pair that $a$ is a common nodes. So we only need to consider case 1.

			Therefore, choose a path $Q$ from input to output with nonzero weight and disjoint from $P,P''$. Now, since weights of $P,P''$ have different signs, and $a$ is in both paths, whatever the weight of $Q$ is, $Q$ and one of  $P$ or $P''$ can form a compensating pair (not weak compensating pair anymore, because $Q$ is disjoint with $P,P''$), and $a$ is in a path. This means $a$ will be opened.

			Combining all discussion above, if $a$ is in a weak compensating path from input to output, then it is in a compensating path from one layer to another layer. Therefore, Theorem \ref{thm:2} implies Theorem \ref{th2}. This is what we want to show.