### 2024-01-25 Paper/Paper_noProof.tex
edits

\documentclass{llncs}
\pagestyle{plain}

%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\vW}{\boldsymbol{W}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vb}{\boldsymbol{bias}}
\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}

\newcommand{\UB}{\mathrm{UB}}
\newcommand{\LB}{\mathrm{LB}}
\newcommand{\ub}{\mathrm{ub}}
\newcommand{\lb}{\mathrm{lb}}
\newcommand{\B}{\mathrm{B}}

\newcommand{\CMP}{{\textrm{CMP}}\ }

\newcommand{\toolname}{\ensuremath{\mathsf{Comp}}}

%\usepackage{amsmath, amsthm, amssymb, amsfonts}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{corollary}{Corollary}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}

\newcommand{\ReLU}{\mathrm{ReLU}}

\title{I Compensate, therefore I Am \\ (accurate for DNN verification)}
\date{}

\begin{document}

	\maketitle

	\begin{abstract}
		Deep Neural Networks (DNNs) verification is now a mature research field, with many methodologies and tools to verify formally that they are correct, with an annual competition to compare them, etc. Formally, the question is, given a DNN and a property to check, does the property hold over a set of inputs of the DNN. This allows e.g. to check local robustness around an input $\vx$, by checking that the i-th output neuron has the largest weight among all output neurons uniformly over a neighbourhood around $\vx$. In the most recent years, the focus has been on combining several of these efficient techniques (branch and bound, multi neuron encoding, MILP encoding, ...) to optimize speed/accuracy trade-offs. While some relatively large DNNs (tens of thousands of neurons) can be checked very efficiently with modern verifiers, some DNNs are still challenging to address with efficient algorithms, even small ReLU DNNs with hundreds of neurons, showing the need for new methodologies.

		In this paper, we analyse efficient algorithms to verify ReLU DNNs, based on abstractions (DeepPoly, Linear Programming, PRIMA, different version of Crown), and uncover the main reason for the loss of accuracy, namely {\em compensations}. Intuitively, a compensation happens when there are 2 paths between the same pair of neurons, one path with positive and one path with negative (product of) weights, which therefore (partly) compensate each other. It is however hard to compute exactly by how much is the compensation due to the ReLU activation functions. Based on this finding, we propose a novel methodology to obtain interesting trade-offs in terms of speed/accuracy, using many queries with few 'open' ReLU nodes (for which both linear modes are considered), rather than the usual few queries with potentially many 'open' ReLU nodes. These promising results open up many different applications of the concept of compensating paths in the field of verification of DNNs.  
	\end{abstract}

	\section{Introduction}

	\input{introduction}

	\section{Background}

	In this paper, we will use lower case latin $a$ for scalars, bold for vectors $\boldsymbol{z}$, 
	capitalized bold for matrices $\boldsymbol{W}$, similar to notations in \cite{prima,crown}.
	To simplify the notations, we restrict the presentation to feed-forward, 
	fully connected ReLU Deep Neural Networks (DNN for short), where the ReLU function is $ReLU : \mathbb{R} \rightarrow \mathbb{R}$ with
	$ReLU(x)=x$ for $x \geq 0$ and $ReLU(x)=0$ for $x \leq 0$, which we extend componentwise on vectors.

	%In this paper, we will not use tensors with a dimension higher than matrices: those will be flattened.

	%\subsection{Neural Network and Verification}

	% testtesttesttest
	An $\ell$-layer DNN is provided by $\ell$ weight matrices 
	$\boldsymbol{W}^i \in \mathbb{R}^{d_i\times d_{i-1}}$
	and $\ell$ bias vectors $\vb^i \in \mathbb{R}^{d_i}$, for $i=1, \ldots, \ell$.
	We call $d_i$ the number of neurons of hidden layer $i \in \{1, \ldots, \ell-1\}$,
	$d_0$ the input dimension, and $d_\ell$ the output dimension.

	Given an input vector $\boldsymbol{z}^0 \in \mathbb{R}^{d_0}$, 
	denoting $\hat{\boldsymbol{z}}^{0}={\boldsymbol{z}}^0$, we define inductively the value vectors $\boldsymbol{z}^i,\hat{\vz}^i$ at layer $1 \leq i \leq \ell$ with
	\begin{align*}
		{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{i-1}+ \vb^i\\
		\hat{\boldsymbol{z}}^{i} &= ReLU({\boldsymbol{z}}^i).
	\end{align*}

	The vector $\hat{\boldsymbol{z}}$ is called post-activation values, 
	$\boldsymbol{z}$ is called pre-activation values, 
	and $\boldsymbol{z}^{i}_j$ is used to call the $j$-th neuron in the $i$-th layer. 
	For $\boldsymbol{x}=\vz^0$ the (vector of) input, we denote by $f(\boldsymbol{x})=\vz^\ell$ the output. Finally, pre- and post-activation neurons are called \emph{nodes}, and when we refer to a specific node/neuron, we use $a,b,c,d,n$ to denote them, and $W_{a,b} \in \mathbb{R}$ to denote the weight from neuron $a$ to $b$. Similarly, for input $\boldsymbol{x}$, we denote by $\val_{\boldsymbol{x}}(a)$ the value of neuron $a$ when the input is $\boldsymbol{x}$. A path $\pi$ is a sequence $\pi=(a_i)_{k \leq  i \leq k'}$ of neurons in consecutive layers, and the weight of $\pi$ is 
	$weight(\pi)=W_{a_k,a_{k+1}} \times \cdots \times  W_{a_{k'-1},a_{k'}}$.

	\iffalse
	and the $i$-th hidden layer is a vector in $\mathbb{R}^{d_i}$, 
	and the output layer is a vector in $\mathbb{R}^{d'}$ or a scale. 
	The weights, bias and activation functions decide propagate the from previous to the next layer. In formula, from layer $l_{i-1}$ to layer $l_{i}$, the weight 
	$\boldsymbol{W}^i$ is matrix of $d_i\times d_{i-1}$, 
	the bias is a vector $\vb^i$ in $\mathbb{R}^{d_i}$, and the activation function 
	is $\sigma$, then  if the $i-1$-th layer is $\hat{\boldsymbol{z}}^{(i-1)}$, 
	then the value of $i$-th layer is computed by: 
	\begin{align*}
		{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \vb^i\\
		\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)).
	\end{align*} The vector $\hat{\boldsymbol{z}}$ is called post-activation values, and $\boldsymbol{z}$ is called pre-activation values, and $\boldsymbol{z}^{(i)}_j$ is used to call the $j$-th neuron in the $i$-th layer. In our style, we also call neurons \emph{nodes} and use $a,b,c,d$ to denote them. We use $W_{ab}$ to denote the weight from neuron $b$ to $a$. We use $\boldsymbol{x}$ to denote the vector of input and  $f(\boldsymbol{x})$ to denote the output.
	\fi

	\medskip

	Concerning the verification problem, we focus for simplicity on the usual local robustness question. Local robustness asks to determine whether the output of a neural network will be affected under small perturbations to the input. 
	Formally, for an input $\vx$ perturbed by $\varepsilon >0$ under distance $d$, then the DNN is locally $\varepsilon$-robust in $\vx$ whenever:
	\begin{align*}
		\forall \boldsymbol{x'} \text{ s.t. } d(\vx,\vx')\leq \varepsilon, \text{ we have }  
		argmax_i (f(\boldsymbol{x'})[i]) = argmax_i(f(\boldsymbol{x})[i])
	\end{align*}

	\iffalse
	In some cases, the output is a vector but the aim to get the label of dimension with the minimal value. In this case, the problem can be written as:\begin{align*}
		\forall \boldsymbol{x} \in\mathcal{D} \  \min f(\boldsymbol{x}) = \min f(\boldsymbol{x}_0)
	\end{align*}

	If so, the question of verification can turn to the following optimization question: \begin{align*}
		\min f(\boldsymbol{x}) \ s.t. {\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ b^i\\
		\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)), \boldsymbol{x}\in\mathcal{D}.
	\end{align*}

	In this paper, we only consider $\ReLU$ function as the activation function: $\sigma(a)=\ReLU(a)=\max(0,a)$.

	In this paper, we consider $L^{\infty}$ norm the max value of distance of each dimension, that is $d(\vx,\boldsymbol{x}_0)=\max |\boldsymbol{x}(n)-\boldsymbol{x}_0(n)|$. 
	\fi

	\section{Value Abstraction for DNN verification}

	In this section, we describe different value (over-)abstractions on $\vz$ that are used by efficient algorithms to certify robustness around an input $\vx$. Over-abstractions of values include all values for $\vz$ in the neighbourhood of $\vx$, and thus a certificate for safety in the over-abstraction is a proof of safety for the original input $\vx$.

	\subsection{The Box and DeepPoly Abstractions}

	%\begin{figure}[t!]
	%	\centering
	%		\includegraphics[scale=0.4]{example.png}
	%	\caption{A DNN example from \cite{kneuron}, where every node is separated into its linear and its ReLU functions.}
	%	\label{fig1}
	%	\end{figure}

\begin{figure}[t!]
	\centering
	\begin{tikzpicture}

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (input1) {$n_1$};

		\node[circle, draw= purple, thick, minimum width = 20,
		minimum height = 20] (input2) at ($(input1) + (0,-1.5)$) {$n_2$};

		% Hidden layers

		\node (hidden10) at ($(input1) + (2.5,0.6)$) {$0$};

		\node (hidden20) at ($(input1) + (2.5,-1.5-0.6)$) {$0$};

		\node (hidden50) at ($(input1) + (7.5,0.6)$) {$0$};

		\node (hidden60) at ($(input1) + (7.5,-1.5-0.6)$) {$1.5$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (hidden1) at ($(input1) + (2.5,0)$) {$n_3$};
		\node[circle, draw= black, thick] (hidden2) at ($(input1) + (2.5,-1.5)$) {$n_4$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (hidden3) at ($(input1) + (5,0)$){$\hat{n}_3$};
		\node[circle, draw= black, thick] (hidden4) at ($(input1) + (5,-1.5)$) {$\hat{n}_4$};

		\node[circle, draw= blue, thick, minimum width = 20,
		minimum height = 20] (hidden5) at ($(input1) + (7.5,0)$){$n_5$};
		\node[circle, draw= black, thick] (hidden6) at ($(input1) + (7.5,-1.5)$) {$n_6$};

		% Output layer
		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (output1) at ($(input1) + (10,0)$){$\hat{n}_5$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (output2) at ($(input1) + (10,-1.5)$){$\hat{n}_{6}$};

		% Connections

		\draw[->,thick] ($(input1) + (-1.5,0)$) -- (input1) node[midway, above] {$[-1,1]$};

		\draw[->,thick] ($(input1) + (-1.5,-1.5)$) -- (input2) node[midway, above] {$[-1,1]$};

		\draw[->,thick] (input1) -- (hidden1) node[near start, above] {$1$};
		\draw[->,thick] (input1) -- (hidden2)node[near start, above] {$1$};

		\draw[->,color=green, thick] (input2) -- (hidden1) node[near start, below] {$1$};
		\draw[->,color=red, thick] (input2) -- (hidden2)node[near start, below] {$-1$};

		\draw[->,color=green, thick] (hidden1) -- (hidden3) node[midway, above] {$\max(n_1,0)$};
		\draw[->,color=red, thick] (hidden2) -- (hidden4) node[midway, above] {$\max(n_2,0)$};

		\draw[->,color=green, thick] (hidden3) -- (hidden5) node[near start, above] {$1$};			
		\draw[->,thick] (hidden3) -- (hidden6) node[near start, above] {$0$};

		\draw[->,color=red, thick] (hidden4) -- (hidden5)node[near start, below] {$2$};
		\draw[->,thick] (hidden4) -- (hidden6)node[near start, below] {$1$};

		\draw[->,thick] (hidden5) -- (output1) node[midway, above] {$\max(n_5,0)$};
		\draw[->,thick] (hidden6) -- (output2) node[midway, above] {$\max(n_6,0)$};

	\end{tikzpicture}
	\caption{A DNN example from \cite{kpoly}: every neuron is separated into 2 nodes, $n$ pre- and $\hat{n}$ post-ReLU activation function. The pair $(n_2 n_3 n_5,n_2 n_4 n_5)$ in green and red is compensating (weights of paths are $1,-2$).}
	\label{fig1}
\end{figure}

The idea to perform value abstraction is to compute upper and lower bounds on the values of (some) neurons in the DNN when the inputs are in a given neighbourhood, in order to conclude on the robustness without needing to compute the values exactly for all inputs (in the neighbourhood).

First, remark that weighted sums is a linear function, so it is easy to represent it explicitly. The only problematic part to represent accurately is the ReLU function. While it is fairly simple piecewise linear function with 2 modes ($x<0$ and $x \geq 0$), it is not linear. Considering its iterated impact along the different layers of a ReLU DNN is much more complex than it seems. First, let us remark that it is easy to represent exactly $\ReLU(x)$ in case $x$ is {\em stable}, that is, either it is known to be entirely positive, order
it is known to be entirely negative, as in both case there is only one known linear mode involved. Hence, the sole problem remains how to handle {\em unstable} neurons.

Consider the simpler abstraction, called the {\em Box abstraction} \cite{deeppoly}: inductively compute the bound on each neuron of the next layer independently, by considering the weighted sum over the bounds of the previous layer, then clip the lower bound at 
$\max(0,$ lower bound$)$ to represent the ReLU function, and so on.
For all $i$, let $x_i=\val_{\vx}(n_i)$, with $\vx=(x_1,x_2)$.
For instance, on the DNN example of Fig \ref{fig1}, $x_1,x_2 \in [-1,1]$, then $x_3,x_4 \in [-2,2]$, using ReLU $\hat{x}_3,\hat{x}_4 \in [0,2]$, and $x_5 \in [0,6]$ while $x_6 \in [0,2]$. 
While the bounds for $n_1, \ldots, n_4$ are exact (for all $\alpha$ 
in the range, there is an input $\vy$ such that $\val_{\vy}(n_i)=\alpha$), 
it is no more the case from the next layer (starting in $n_5,n_6$) because of potential 
dependencies between earlier neurons: $n_3$ reaches value $2$ only when $x_1=x_2=2$,
in which case $n_4$ reaches value $Val_{(2,2)}(n_4)=0$. Hence, it is impossible that $x_5=\Val_{\vx}(n_5)$ reaches $6$, as it would require both $n_3$ and $n_4$ to have value $2$.

A still extremely fast algorithm that corrects some of this issue is {\em DeepPoly} \cite{deeppoly}, also rediscovered as the {\em CROWN} algorithm \cite{crown}. Instead of 
hard bounds, for each neuron $n$ of layer $k$, two affine functions (with input at the previous layer $k-1$) are kept as lower bound and upper bound of the value of the node.
Denote $f_i \leq x_i \leq g_i$ with e.g.  
$f_{3}(x_3)=f_4(x_4)=0$ and 
$g_3(x_3) = \frac{x_3+2}{2}$,
$g_4(x_4) = \frac{x_4+2}{2}$. 
With that, we get 
$x_5 \leq g_3(x_3) + 2 g_4(x_4) = \frac{x_3 + 2x_4 + 6}{2} = \frac{3x_1 - x_2 + 6}{2}\leq 5$.

For $[\alpha,\beta]$ bounds on $x_i$, there is one optimal linear function for the upper bound $g_i(x_i)= \beta \frac{x_i-\alpha}{\beta-\alpha}$ for ReLU nodes.
There are two choices for the lower bound: $f^1_i(x_i) = 0$ or $f^2_i(x_i)=x_i$.
In the DeepPoly algorithm, the choice is made by considering $\alpha < 0 < \beta$ (for unstable neurons). If $|\alpha|\geq |\beta|$ then $f_i=f^1_i$ is used, while if $|\beta|>|\alpha|$ then $f_i=f^2_i$ is used. Denote {\em $\overline{\mbox{DeepPoly}}$} the DeepPoly algorithm such that $f^1_i$ is always used, and $f^2_i$ is never used. 
Unlike DeepPoly, {\em $\overline{\mbox{DeepPoly}}$} subsumes the {\em Box abstraction}. Indeed, taking bounds $[-0.2,5]$ on $x_i$, it gives $\ReLU(x_i) \in [0,5]$, 
while the normal DeepPoly algorithm would only allow to conclude that $\ReLU(x_i) \in [-0.2,5]$.

\iffalse
\subsection{PRIMA and $\beta$-CROWN}
\fi

\subsection{MILP and LP encodings for DNNs}

At the other end of the spectrum, we find the Mixed Integer Linear Programming (MILP) value abstraction, which is a priori a complete method (albeit much slower than DeepPoly). 
Consider an unstable neuron $n \in[\alpha,\beta]$. The value $x$ of $\ReLU(n)$ can be encoded exactly in an MILP formula with one integer (actually even binary) variable $a$ valued in ${0,1}$, using 4 constraints \cite{MILP}:
\vspace{-0.1cm}
\begin{align*}
	\hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0, \quad \wedge \quad \hat{x} \leq \beta \cdot a \quad \wedge \quad \hat{x} \leq x-\alpha \cdot (1-a)
\end{align*}

From \cite{MILP}, for all $x \in [\alpha,\beta] \setminus 0$, there is a unique solution $(a,\hat{x})$ satisfying the constraints, and we have $\hat{x}=\ReLU(x)$ (and $a$ is 0 if $x < 0$ and 1 if $x>0$, and can be either for $x=0$). Such an encoding can be used for every (unstable) ReLU node, and optimizing its value allows to certify a given input. 
For Networks with hundreds of nodes or more, the associated MILP formula will have have hundreds of integer variables, and it will usually not be solvable efficiently.

MILP instances can be linearly relaxed into LP over-abstraction, where variables which were integers in $\{0,1\}$ (binary) are relaxed to be real in interval $[0,1]$, using otherwise the same encoding. Solving LP instances being Polynomial time, optimizing it is much more efficient, but at the price of a lack of accuracy (e.g. with not as tight bounds). This is called the {\em LP abstraction}.

As far as we know, no formal result states the relationship between the LP approximation and DeepPoly.
We prove now that the LP abstraction is strictly more accurate than the DeepPoly abstraction, as it actually corresponds to considering in DeepPoly both linear functions lower bounding the ReLU function, $ReLU(x) \geq 0$ and $ReLU(x) \geq x$, while DeepPoly considers exactly one.

\begin{proposition}
	\label{LP}
	Given $x \in [\alpha,\beta]$ with $\alpha < 0 < \beta$, the following two systems of constraints 
	1) for LP and 2) for DeepPoly with both lower bounds are equivalent:
	\vspace{-0.3cm}
	\begin{align*}
		& 1) \quad \hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0, \quad \wedge \quad \hat{x} \leq \beta \cdot a \quad \wedge \quad \hat{x} \leq x-\alpha \cdot (1-a), \, a \in [0,1] \\
		\text{and} \quad  & 2)  \quad \hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0 \quad \wedge \quad \hat{x} \leq \beta \frac{x-\alpha}{\beta-\alpha}
	\end{align*} 
\end{proposition}

\begin{proof}
	Obviously, the 2 lower bound constraints are the same in 1) and 2).

	It remains to prove that the 2 upper bound constraints of 1) LP corresponds to the 2) DeepPoly linear upper bound. For that, we study the upper bound function in the linear variable $a \in  [0,1]$, with $x \in [\alpha,\beta]$ fixed. We have $\hat{x}$ in 1) is upper bounded by $max_{a \in [0,1]} (min(\beta \cdot a, x - \alpha (1-a)))$, and this bound can be reached. 
	The function $min(\beta \cdot a, x - \alpha (1-a))$ reaches its maximum for 
	$\beta \cdot a = x - \alpha (1-a)$, that is 
	$(\beta - \alpha) a = x - \alpha$, i.e. 
	$ a = \frac{x - \alpha}{\beta-\alpha}$.
	It thus gives an upper bound at 
	$\beta \cdot a = \beta \frac{x - \alpha}{\beta-\alpha}$, which is exactly the upper-bound function used in DeepPoly.
	\hfill $\square$
\end{proof}

\section{Compensating pairs of paths}
\label{Sec.comp}

In this section, we explore the key factor for the loss of accuracy in value abstraction methodologies, as those we presented before (except for full MILP which is exact but 
not efficient). 
Namely, we uncover that \emph{compensating pair of paths} are the reason for the loss of precision. A pair $(\pi,\pi')$ of paths is {\em compensating} if:
\begin{itemize}
	\item they have the same starting node $a$ (called {\em source} state) and ending node $b$ (called {\em target} state of $(\pi,\pi')$),
	\item they are disjoints (the only common nodes are the source and the target),
	\item the weights satisfy $weight(\pi') < 0 < weight(\pi)$.
\end{itemize}

For instance, on Fig.~\ref{fig1}, the paths $\pi= n_2 n_3 n_5$ has weight $1$, while the
path $\pi'= n_2 n_4 n_5$ has weight $-2$, hence $(\pi,\pi')$ is a compensating pair of paths.
We call a path $\pi$ {\em compensating} if there exists a path $\pi'$ such that either $(\pi,\pi')$ or $(\pi',\pi)$ is compensating.

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $x_5 \leq 4$, which is reached for $x_1=1$ and $x_2=-1$: 
$\val_{(1,-1)}(n_5)=4$. However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (contribution of $\frac{-x_2+3}{2} \leq 2$ instead of exact $-x_2 \leq 1$), obtaining $x_5 \leq 5$ instead of the precise $x_5 \leq 4$.

Intuitively, the target state receives the weight from the source state from the different paths from the source to the target. As the weights of the paths is of opposite sign, some of these weights will cancel out, by an amount up to $min(|weight(\pi')|,|weight(\pi')|)$. 
Further, that also avoid accumulating the error, in the same way as $0= |1+(-1)| \leq |1|+|-1|=2$ from an inaccurate analysis. The amount of compensation can however be reduced by the ReLU functions, because of clipping by the ReLU function. For instance, consider that the inputs $x_1,x_2$ lies in $[-1,-0]$. Then $\hat{x}_3$ will always be 0 for any input $\vx=(x_1,x_2) \in [-1,-0] \times [-1,-0]$, and $weight(\pi')$ will not compensate $weight(\pi')$.

\iffalse
\vspace*{2ex}

\begin{figure}
	\centering
	\begin{tikzpicture}

		\node[circle, draw= purple, thick, minimum width = 20,
		minimum height = 20] (input1) {$a$};

		% Hidden layers
		\node[circle, draw= blue, thick, minimum width = 20,
		minimum height = 20] (hidden1) at ($(input1) + (2,1)$) {$b$};
		\node[circle, draw= blue, thick] (hidden2) at ($(input1) + (2,-1)$) {$b'$};

		\node[circle, draw= blue, thick, minimum width = 20,
		minimum height = 20] (hidden3) at ($(input1) + (4,1)$){$c$};
		\node[circle, draw= blue, thick] (hidden4) at ($(input1) + (4,-1)$) {$c'$};

		% Output layer
		\node[circle, draw= blue, thick, minimum width = 20,
		minimum height = 20] (output) at ($(input1) + (6,0)$){$d$};

		% Connections
		\draw[->,thick,draw= red] (input1) -- (hidden1) node[midway, below] {$1$};
		\draw[->,thick,draw= green] (input1) -- (hidden2)node[midway, below] {$-1$};

		\draw[->,thick,draw= red] (hidden1) -- (hidden3) node[midway, below] {$\ReLU$};
		\draw[->,thick,draw= green] (hidden2) -- (hidden4) node[midway, below] {$\ReLU$};

		\draw[->,thick,draw= red] (hidden3) -- (output)node[midway, below] {$1$};
		\draw[->,thick,draw= green] (hidden4) -- (output)node[midway, below] {$1$};
	\end{tikzpicture}
\end{figure}

\vspace*{2ex}

In this figure, $a$ is the input neuron; $bc,b'c'$ are nodes in the hidden layer, ($b,b'$ are pre-activation and $c,c'$ are post activation); and $d$ is the unique output neuron. The numbers next to the arrows are the weights. So, $W_{ba}=1$ and $W_{b'a}=-1$, $W_{dc}=W_{dc'}=1$. The pair of these two paths, $a$ to $bc$ to $d$, and $a$ to $b'c'$ to $d$, is a so called \emph{Compensating Pair}. Because its shape looks like a diamond, it is also called a Diamond. The characteristic is that, the products of all weights in the paths, have two different signs: along $bc$, the product is (strictly) positive, while along $b'c'$, the product is (strictly) negative.

The existence of compensating pairs is key reason why simple approximation like LP or Interval Arithmetic cannot get the exact upper and lower bounds. If both pairs are negative or positive, LP or even Interval Arithmetic will get the exact values of lower and upper bounds.

To explain why, suppose we have another input node $a'$, such that both $a$ and $a'$ has an input interval $[0,1]$, but the weight from $a'$ to $b$ or $b'$ are both $1$. Then both $b$ will have an interval $[0,2]$ and $b'$ will have an interval $[-1,1]$. More importantly, in LP formulation, $c$ will be $b$ but $c'$ will be $0.5(b')+0.5$. The upper bound of $d$ will be $c+c'$, and hence $b+0.5(b')+0.5$, and hence $a+a'+0.5a'-0.5a'+0.5=0.5a+1.5a'+0.5$. And this will leads to a upper bound $2.5$ but this is not exact.

\vspace*{2ex}
\begin{tikzpicture}
	\node[circle, draw= purple, thick, minimum width = 20,
	minimum height = 20] (input1) {$a$};

	\node[circle, draw= purple, thick, minimum width = 20,
	minimum height = 20] (input2) at ($(input1) + (0,-2)$) {$a'$};

	% Hidden layers
	\node[circle, draw= blue, thick, minimum width = 20,
	minimum height = 20] (hidden1) at ($(input1) + (2,0)$) {$b$};
	\node[circle, draw= blue, thick] (hidden2) at ($(input1) + (2,-2)$) {$b'$};

	\node[circle, draw= blue, thick, minimum width = 20,
	minimum height = 20] (hidden3) at ($(input1) + (4,0)$){$c$};
	\node[circle, draw= blue, thick] (hidden4) at ($(input1) + (4,-2)$) {$c'$};

	% Output layer
	\node[circle, draw= blue, thick, minimum width = 20,
	minimum height = 20] (output) at ($(input1) + (6,-1)$){$d$};

	% Connections
	\draw[->,thick,draw= red] (input1) -- (hidden1);
	\draw[->,thick,draw= green] (input1) -- (hidden2);

	\draw[->,thick,draw= red] (input2) -- (hidden1) node[midway, below] {$1$};
	\draw[->,thick,draw= red] (input2) -- (hidden2)node[midway, below] {$1$};

	\draw[->,thick,draw= red] (hidden1) -- (hidden3) node[midway, below] {$\ReLU$};
	\draw[->,thick,draw= green] (hidden2) -- (hidden4) node[midway, below] {$\ReLU$};

	\draw[->,thick,draw= red] (hidden3) -- (output)node[midway, below] {$1$};
	\draw[->,thick,draw= green] (hidden4) -- (output)node[midway, below] {$1$};
\end{tikzpicture}
\vspace*{2ex}

The general formal definition of compensating pair is as follows:

\begin{definition} In a full-connected network with $\ReLU$ as activation function:

	1. A path is a sequence of nodes $\langle a,b,c,d,e,\cdots\rangle$ of nodes that goes consecutively through each layer. We call the first node source node and the last node target node.

	2. The \emph{Value} of a path is the product of of weights along the path (with sign): for a path $\langle a,b,c,d,e,\cdots\rangle$, its values is $$V = W_{ab}\cdot W_{bc}\cdot W_{cd}\cdot W_{ed}\cdot \cdots$$

	3. A \emph{Compensating Pair} is a pair of paths with the same source node and target node, such that the two paths have no common node, and the values of two paths have opposite signs (one is strictly positive and another is strictly negative).

	We also use \emph{Diamond} to call a compensating pair in the network.
\end{definition}

The following theorem shows the role of compensating pairs in the computation:

\fi

Hence, compensation is {\em one} factor of inaccuracy. We prove now that it is actually the {\em only} reason for inaccuracy, in the following meaning.
If there is no compensating path in the DNN, then any method/approximation at least as accurate as the Box abstraction, e.g. $\overline{\text{DeepPoly}}$ or LP, computes the exact upper and lower bounds for every node, that is:

\begin{table}[b!]
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		\text{Source/Target Layers}  &  \text{Natural DNN} & \text{Robust DNN} & \text{Ratio Natural vs Robust} \\ \hline \hline
		0 / 2 & 0.0304 & 0.00220  & 13.8x\\ \hline
		1 / 3  & 0.0313 & 0.00875 & 3.58x \\ \hline
		2 / 4  &  0.0267 & 0.00785 & 3.40x \\ \hline
		3 / 5  &  0.0253 & 0.00804  & 3.18x \\ \hline
	\end{tabular}
	\caption{Comparison of the average compensation strength over all the pairs of nodes of layer source/target between a DNN naturally-trained and a DNN robustly-trained using DiffAI \cite{DiffAI} with the same architecture and training set.}
	\label{tab:compensation}
\end{table}

\begin{theorem}
	\label{th1}
	Let $n$ be any node of a DNN with {\em no} compensating path. For $[\alpha,\beta]$ the bounds computed by the {\em Box abstraction} for $n$, we have that for all $\gamma \in [\alpha,\beta]$, there exists an input $\vx$ such that $\val_{\vx}(n)=\gamma$.
\end{theorem}

The sketch of proof can be found in Section \ref{sec.proofs}. 
Notice that standard DeepPoly does not satisfies Theorem \ref{th1} (it does not necessarily refine Box), and therefore we are not sure PRIMA or $\beta$-CROWN that refine DeepPoly satisfies Theorem \ref{th1}, though it would be easy to fix by using $\overline{\text{DeepPoly}}$. Anyway, in practice, it is very unlikely to have a network without any compensating path, it is more of a theoretical results show in that compensation is the reason for inaccuracy.

This result has a first interesting application: explain why some DNNs are intrinsically easier to verify than others. This is the case for DNNs trained adversarially to be robust, vs same architecture, but trained in a "natural" way \cite{deeppoly,prima,crown}. Indeed, compensating pairs is a purely structural property on the DNN learnt, rather than more semantical notions, such as the number of unstable ReLUs, which depends heavily upon the image considered and the algorithm used. We compare compensations with source and target separated by 1 ReLU layer, e.g. source in layer 0 and target in layer 2, for two Networks with 5 hidden layers of 100 nodes each from the ERAN GitHub repository, one trained naturally ("$6\times100$"), 
and one trained using DiffAI \cite{DiffAI} ("$5\times100$"). 
We report in Table \ref{tab:compensation} the average compensation strength, i.e. 
the average of $max_{\pi,\pi'} min(|weight(\pi')|,|weight(\pi')|)$
over all source/target states in given layers. The average compensations strength are much lower in the Robust DNN than in the natural DNN, and indeed the Robust DNN is much simpler to verify than the natural DNN (e.g. DeepPoly is more accurate and verifies more images).

\subsection{MILP$_{X_n}$}

While Theorem \ref{th1} is interesting to understand that compensation is a key notion for accurate verification of DNNs, it cannot be used directly to analyze DNNs with compensations, which are actually those that are most interesting to tackle.
Intuitively, when there are compensating paths, it seems necessary to consider exactly the (unstable) ReLU nodes that are seen along these paths. 
Consider a neuron $n$ in layer $k$ for which we want to have accurate bounds 
$[\alpha,\beta]$.
Let $X_n$ be the set of neurons $x$ in layer up to $k-1$ such that there is a compensating path with target $n$ passing by $x$ (but not as its source or target). Let $Y_n$ be the complement of this set $X_n$ of neurons in the first $k-1$ layers. We denote by MILP$_{X_n}$ the relaxation of the MILP encoding where all the nodes in $Y_n$ are linearly relaxed (thus with $|X_n|$ binary variable, one per neuron in $X_n$), and including exact bounds for every neurons $n'$ in the first $k-1$ layers. We state now that the bounds $[\alpha,\beta]$ on neuron $n$ computed by such MILP$_{X_n}$ are accurate.

\begin{theorem}
	\label{th2} 
	Let $n$ be any node of a DNN. Then for $[\alpha,\beta]$ the bounds computed by MILP$_{X_n}$ for neuron $n$, for all $\gamma \in [\alpha,\beta]$, there exists an input $\vy$ such that $\val_{\vy}(n)=\gamma$.
\end{theorem}

The sketch of proof of Theorem \ref{th2} can be found in Section \ref{sec.proofs}.

\subsection{Sketch of Proofs}
\label{sec.proofs}

%Intuitively, our first proof shows that if there are no compensation, then there are no correlations between nodes.
Consider a target neuron $z$.
In case there is no compensating path, we can assign a sign to each neuron $n$: 0
if all paths from $n$ to $z$ has weight 0, $+1$ if all paths have positive weights, and 
$-1$ if they have negative weights. Then we prove:

\begin{proposition}
	\label{prop.sign}
	There exist two inputs $\vx^\star,\vx^\sharp$ such that ${\vx^\star}$ maximizes the value of all positive neurons and minimizes the value of all negative neurons, and  ${\vx^\sharp}$ minimizes the value of all positive neurons and maximizes the value of all negative neurons.
\end{proposition}

Hence $\val_{\vx^\star}(z)=\beta$ and $\val_{\vx^\sharp}(z)=\alpha$,
for $[\alpha, \beta]$ the bounds computed by the Box abstraction. We conclude Theorem \ref{th1} by invocking continuity.

\smallskip

Concerning Theorem \ref{th2}, consider an output node $z$.
Let $[\alpha(z),\beta(z)]$ be the bound computed by $\MILP_{X_z}$.
%We will show the existence of $\vx^\sharp,\vx^\star$
%such that $\val_{\vx^\sharp}(z)=\alpha$ and $\val_{\vx^\star}(z)=\beta$,
%and we will get the proof of Theorem 2 by a continuity argument.
%For that, 
We partition the input neurons from layer $0$ into:
\begin{enumerate}
	\item $A_{zero}= \{a \mid \forall \text{ path $\rho$ from $a$ to } z, weight(\rho)=0\}$.
	\item $A_{pos}= \{a \mid \forall \text{ path $\rho$ from $a$ to } z, weight(\rho)\geq0\}$.
	\item  $A_{neg}= \{a \mid \forall \text{ path $\rho$ from $a$ to } z, weight(\rho)\leq0\}$.
	Let $A_{pure}=A_{pos} \cup A_{neg}$.
	\item $A_{open}$ is the set of remaining input neurons.
\end{enumerate}

We then partition the set of neurons in hidden layers: 
\begin{enumerate}
	\item $B_{zero}= \{n \mid \forall \text{ path $\rho$ from $n$ to } z, weight(\rho)=0\}$.
	\item $B_{open}$ is the set of neurones reachable with $>0$ weight from $A_{open}$ and such that there is at least one path to $z$ with non zero weight.
	\item $B_{pure}$ is the set of remaining neurons.
\end{enumerate}

We can define the sign function for neurons in $A_{pure}$ and $B_{pure}$.
Further, the value of any neuron $n$ in $A_{pure}$ or $B_{pure}$ is entirely determined by 
nodes in $A_{pure}$ or $B_{pure}$. Therefore, applying Prop. \ref{prop.sign} on the DNN
made of nodes from $A_{pure}$ and $B_{pure}$, we obtain 
$\vx_{pure}^\star,\vx_{pure}^\sharp$ optimizing neurons in $A_{pure}$ or $B_{pure}$.
We can prove that there exists input 
$\vx^\star$ extending $\vx_{pure}^\star$
and input $\vx_{pure}^\sharp$ extending $\vx_{pure}^\sharp$
from $A_{pure}$ to $A_{pure} \cup A_{open}$ such that 
$\val_{\vx^\star}(z)=\beta(z)$ and $\val_{\vx^\sharp}(z)=\alpha(z)$.

The formal proofs are in appendix.

\subsection{An efficient algorithm}

Using Theorem \ref{th2}, one could thus obtain accurate bounds by running MILP$_{X}$ inductively on every node layer by layer, obtaining accurate bounds $[\alpha,\beta]$ that can be used to compute the next layer accurate bound. In practice, such an algorithm would however be inefficient, as $X_n$ comprises most of the nodes from layer up to $k-1$ for $n$ on layer $k$. Instead, an interesting trade off between speed and accuracy is to set a threshold on the the strength of compensation that are meaningful, and consider that paths below this threshold are not compensating strongly enough to deserve an integer variable, calling $Z_n$ the resulting set of nodes on {\em strongly compensating} paths. In practice, one could take the $K$ nodes covering the strongest compensating pair of paths to obtain set $Z_n$, for $K$ reasonably small. The pseudo-code of the algorithm is provided in Algorithm \ref{algo1}.

\SetKwInput{KwInput}{Input}
\SetKwInput{KwOutput}{Output}

\begin{algorithm}[b!]
	\caption{Compensate(K)}
	\label{algo1}
	\KwInput{Bounds $[\alpha_n,\beta_n]$ for input nodes $n$ at layer $0$ (input neighbourhood)}

	\KwOutput{Bounds $[\alpha_n,\beta_n]$ for every node $n$}

	\For{layer $k=1 \cdots \ell$}{
		\For{neuron $n$ in layer $k$}{

			Compute $Z$ a set of $K$ nodes covering the compensating pairs of paths with target $n$
			with heaviest compensation

			Run MILP$_Z$ to obtain $[\alpha_n,\beta_n]$ from bounds of neurons in layers $< k$
		}
	}
\end{algorithm}

This algorithm has a worst case complexity bounded by $O(N \MILP(N,K))$, 
where $N$ is the number of nodes of the DNN, 
and $\MILP(N,K)$ is the complexity of solving a MILP program with $K$ integer variables and $N$ linear variables.
We have $\MILP(N,K) \leq 2^K \LP(N)$ where $\LP(N)$ is the Polynomial time to solve a Linear Program with $N$ variables.
This complexity is an upper bound, as e.g. Gurobi is fairly efficient and never need to consider all of the $2^K$ ReLU configurations to compute the bounds.
Notice that the for loop on neurons $n$ in layer $k$ at line 2 can be realized in parallel as we only need bounds from previous layers, not from the current layer $k$, 
leading to a time complexity of $O(\sum_{i=1}^{\ell} \MILP(N_{i-1},K))$, with $\ell$ the number of layers and $N_i$ the number of neurons in layers $0, \ldots, i$.
Hence, for $K$ small enough, Algorithm \ref{algo1} calls many times MILP (e.g. Gurobi)
with a limited number of integer variables, which should be efficient enough, and yet be quite accurate as it can compute exactly (Theorem \ref{th2}) the highest compensating path, which we verify in the next section.

\iffalse
\section{Verification Framework}

%Our experiments are carried by different version codes, and the global process has been changed.

In this section, we will sketch the framework.

\subsection{Structure}

\subsubsection*{Precomputation}

The open node chosen is the key part of the whole framework but costs a lot of time. So some computation (those do not rely on certain image and bounds) is moved to the precomputation part.

%The precomputation corresponds to the simplest case: source node fixed. We will compute and store a limited number of paths with highest (absolute) values before running any image. This does not rely on other parameters or results(bounds).
%
%Before build an MILP model and when choosing open nodes list,  the program will read the reference data combining with the data of bounds to continue the computation of open nodes. This will use much less time.

\subsubsection*{Process for one image} An image is the basic unit in whole process. The process of one image is simple: compute the concrete bounds of nodes layer by layer, use the bounds of previous layer to build MILP models of current layer. In one layer, nodes runs in parallel.

%Suppose we have reached a new layer $l_i$ and have upper and lower bounds of all nodes in previous layers. Then we will compute bounds of every node in $l_i$ in parallel using the data of bounds of previous nodes. The method is to build and optimize an MILP model.

%Some parameters may be changed during layers. Among all parameters, two groups are the most important: numbers of open nodes and local timeout parameters. Here, \emph{local} is opposite to \emph{global}. We have global timeout parameters for images and the whole running. Local timeout are used in one layer, one node, one model, or one loop in the optimization for a model.

%The open node chosen is the key part of the whole framework, and it will cost two much time without precomputation because we must do open node chosen for every node. 
%
%The precomputation corresponds to the simplest case: source node fixed. We will compute and store a limited number of paths with highest (absolute) values before running any image. This does not rely on other parameters or results(bounds).
%
%Before build an MILP model and when choosing open nodes list,  the program will read the reference data combining with the data of bounds to continue the computation of open nodes. This will use much less time.

\begin{algorithm}
	\caption{The Frame work}
	\KwData{Input Domain: $\mathcal{D}$}
	\KwResult{Bounds of all nodes: $\mathcal{B}$}

	Constant $\mathcal{R}$  \tcp*{Stored precomputation data}

	Constant $\boldsymbol{W}, \vb$ \tcp*{Weights and Bias of network}

	Constant Layers = $[0,1,2\cdots,L]$ \tcp*{lists of all layer}

	Constant NodesLayer = $[Nodes_0,Nodes_1,\cdots, Nodes_L]$ \tcp*{list of nodes in each layer}

	Initialize $\mathcal{B}$ = \{\} \tcp*{To store upper and lower bounds of all nodes}

	\For{$l$ in Layers}{
		\If{$l$ = 0}{

			$\mathcal{D}$, $\boldsymbol{W}, \vb$ $\rightarrow$ $\mathcal{B}_0$ = \{$x$: $(UB(x),LB(x))$,$\cdots$\} 
			\tcp*{Do linear transformation  from input to the first layer}

			Add \{$x$: $(UB(x),LB(x))$,$\cdots$\} to $\mathcal{B}$}

		\Else{

			$l$, NodesLayer $\rightarrow$ $Nodes_l$ = $[x_0,x_1,\cdots]$

			\For{$x$ in $Nodes_l$}{
				$x$, $\mathcal{B}$, $\mathcal{R}$  $\rightarrow$ $\mathcal{O}$ \tcp*{To get the open node list $\mathcal{O}$}

				$\mathcal{B}$, $\boldsymbol{W}, \vb$, $\mathcal{O}$ $\rightarrow \mathcal{M}$  \tcp*{MILP model $\mathcal{M}$ for node $x$}

				\While{}{$\mathcal{M}$ $\rightarrow UB(x),LB(x)$ \tcp*{Optimization}}

				Add \{$x$: $(UB(x),LB(x))$\} to $\mathcal{B}$

			}

		}

	}

	\Return{$\mathcal{B}$}

\end{algorithm}

\subsubsection*{Process for one node}

The optimization of a model consists of loops. For each loop, the model will do optimization by a timeout and observe the result. If it gets improvement, then continue the loop until the optimize bounds or the longer timeout. Otherwise it will give up optimization and store the best bounds obtained so far.

\subsubsection*{Global Process}

In the newest version, we do images by batches for 100 images each. For each batch, the running consists of three turns, very fast turn, fast turn and slow turn.

In the very fast turn, it will run DeepPoly for all images to get a preliminary bounds for all nodes and verify the easiest images. Images verified and images with false predication will be deleted from the image list.

Then sort all remain images by the size of uncertainty of DeepPoly from smaller to larger. The fast turn is to run the images from the smaller side to larger side until consecutive two images cannot be verified. All remain images and images tried but not verified will be put into the next list. And then run the list with parameters for slow turn.

\subsection{Parameters}

Parameters like open node parameter $O$ is important in the frame. Basically, all parameters will be reset after one image.

\subsubsection*{Global parameters}

\subsubsection*{Local parameters}

$O$ and timeout parameters for loops will change frequently during the loops of one node. These change will not reset until the end of one layer or one image.

\subsection{Other Setting}

In our experiments, we have not used PGD-attack to exclude some images. In principle, there is no obstacle to use PGD-attack.
\fi

\input{experiments}

\vspace{-0.2cm}
\section{Conclusion}
\vspace{-0.2cm}

In this paper, we introduced the notion of {\em compensating pairs of paths}, with rationale why such a phenomenon creates inaccuracies hard to handle when verifying DNNs. We proved that this phenomenon is actually explaining entirely the inaccuracies, as in their absence, even the simplest Box abstraction (interval arithmetic) suffices to verify accurately DNNs. We verified experimentally that DNNs learnt in a natural way, harder to verify than robustly trained DNNs, exhibit paths with larger compensating strength.

Based on this idea of compensating pairs of paths, we proposed abstraction $\MILP_{Z}$ considering a subset $Z$ of the set of unstable ReLU nodes, that we proved to be fully accurate if {\em all} the compensating paths are covered by $Z$. Last, we experimentally showed that choosing $Z$ covering {\em few} paths with the strongest compensations, results in very good accuracy, and runtime which can be favourable in some (but not all) cases. This proves that compensating strength is a promising novel concept which should be further studied to improve DNN verification. We finally proposed several ways to use the compensating strength to optimize current tools in different directions, and leave that as future work.

\newpage

\bibliographystyle{plain}
\bibliography{references}

\newpage

\appendix

\input{proofs}

\end{document}

\section{Ranking of Compensating Pairs}

In this subsection, we will introduce the method to choose open nodes for MILP model based on compensating pairs.

Based on No Diamond Theorem, for a target node, in its MILP model, if we open all nodes in compensating pairs, then we can get the exact values of upper and lower bound of the target node. However, in practice, this is still too expensive because we will still need to open too many nodes. So we can only choose a limited number of compensating pairs and nodes. This is by setting a parameter $O$ such that the process only choose at most $O$ many nodes to open. Therefore, in this subsection, we will introduce the process of open node chosen: given a target node, choose up to $O$ many nodes to open.

Basically, we do the process of open nodes chosen for one target node each time, that is, receive one target node (and other data) as input each time. But in principle, we can develop a method receive more than one nodes as input. But in our experiences, this does not work well.

This process is not the unique method but the most natural way to choose open nodes. We introduce it in three steps from the simplest case to more complex case.

\subsubsection*{Value of a pair}

For a compensating pair, if $V_1,V_2$ are the values of two paths, then the value of this compensating pair is defined by: $V=\min(|V_1|,|V_2|)$.

\subsection*{Case: Source node fixed}

The simplest case is when we only consider compensating pairs with a fixed source node.

In this case, the process is very simple: enumerate all compensating pairs, and then sort them by their values. Finally, from the first to the last compensating pair, pick all unstable nodes in its two paths except the target and source nodes into the list of open nodes, until $O$ many open nodes.

%\subsubsection*{Sort all paths by values}
%
%The first step is to sort all paths by their values and divide paths into two groups, a group of paths with positive values and a group of paths with negative values.
%
%Since we have set a bound $O$ for open nodes, we will store  a fixed number of path for each group.
%
%\subsubsection*{Sort pairs by values}
%
%The second step is to sort all pairs by their values. Enumerate paths from the positive group and the negative group one by one and put the pairs obtained into a new list of pairs. Then sort all pair by their values from largest to the smallest: recall that the value of a pair $\langle P_1,P_2\rangle$ is $\min(|V_1|,|V_2|)$.
%
%\subsubsection*{Choosing nodes}
%
%The third step is to choose nodes from the list of pairs. According to the sorted list of pairs, enumerate pair one by one; for each pair, pick the nodes unstable in the two paths except the source and target node into the open node list. Repeat this process until $O$ nodes chosen or reach the end of the list.
%
%\subsubsection*{Pseudocode}
%
%The following needs a chart of pseudo-code
%
%\vspace*{1ex}
%
%1. Enumerate all path from the fixed source node to the target node. 
%
%2. For each path, compute its weight, that is the products of all $W_{aa'}$ along the path.
%
%3. Divide all paths into two group: positive paths and negative paths.
%
%4. Pair positive paths and negative paths from those with larger absolute values to smaller. 
%
%5. For each pair, its value is the min of the weight positive paths and absolute negative values.
%
%6. From pairs with larger values to smaller, pick path one by one, and check all intermediate nodes (nodes except source and target) of each path, and if any of them is unstable ($u>0$ and $l<0$), then open this node. 
%
%7. Repeat 6 until choosing sufficiently many nodes.
%

\subsection*{Case: Source nodes in one layer}

The more general case is when the source node can be any node in a fixed layer. In this case, the process is very similar to previous case, except the value of a pair.

In this case, for a pair $P_1,P_2$ with values $V_1,V_2$ with the source node $a$, its value is $\min(|V_1|,|V_2|)\times \text{upper bound of } a$.

%\subsubsection*{Pseudocode}
%
%The following needs a chart of pseudo-code
%
%\vspace*{1ex}
%
%1. Enumerate all path from the fixed source node to the target node. 
%
%2. For each path, compute its weight, that is the products of all $W_{aa'}$ along the path.
%
%3. Divide all paths into two group: positive paths and negative paths.
%
%4. Pair positive paths and negative paths from those with larger absolute values to smaller. 
%
%5. For each pair, its value is the min of the weight positive paths and absolute negative values, times the upper bound (at least 0) of the source node.
%
%6. From pairs with larger values to smaller, pick path one by one, and check all intermediate nodes (nodes except source and target) of each path, and if any of them is unstable ($u>0$ and $l<0$), then open this node. 
%
%7. Repeat 6 until choosing sufficiently many nodes.

\subsection*{Case: Source nodes in different layers}

The general case is that when the location of source nodes can be in different layers. This case is much more complex, because the scale level of values of paths from different layers are very different: usually a weight will be very small since every layer has a large number of nodes, therefore values of longer paths are products of more small numbers, and is definitely much smaller. If we sort all compensating pairs from different layers directly, then a very long initial segment of this list will be occupied by compensating pairs with the shortest length. Therefore, we need use some scale factors when comparing values of compensating pairs of different length.

In our experiments, we only consider the paths of length 3 (the source node is 2 layers before) and length 4 (the source node is 3 layers before). We will dynamically adjust the values during the process of node chosen. In text, the process of node chosen for length 3 plus length 4 is as follows:

\vspace*{1ex}

1. Generate the two lists of pairs sorted by their values for source nodes in 2 layers before the target node and 3 layers before separately as in the previous case.

2. Enumerate pair from two lists one by one by their values and pick nodes as previous case. When comparing the values between pairs of length 4 and pairs of length 3, multiply the numbers of nodes in 1 layer before to the value of pair of length 4.

In formula, when the target node is in layer $L$, for a pair $P_3$ of length 3 with value $V_3$ and a pair $P_4$ of length 4 with value $V_4$, if we have chosen $N$ nodes in layer $L-1$. Then the adjusted values for $P_3$ and $P_4$ are: $$V_3, N\cdot V_4.$$ Then we pick the next pair by the adjust value in two lists.

3. Repeat 2 until choosing sufficiently many nodes.

\vspace*{1ex}

The pseudo-code is as follows ():

\begin{algorithm}
\caption{Process for length 3 + length 4}
\KwData{Sorted lists of length 3 + length 4}
\KwResult{Open node list}

Let $N = 0$ be the number of Chosen length 4 pair.

\While{len(Open node) $<O$}{
	Pick the first element $P_3$ in the list of length 3\;

	Let $V_3$ be the value of $P_3$\;

	Pick the first element $P_4$ in the list of length 4\;

	Let $V_4$ be the value of $P_4$\;

	\If{$V_3 > V_4 \cdot N$}{
		Add nodes in $P_3$ into Open list\;
		Number of Chosen lentgh pair $N$ += 1\; 
	}
	\Else{
		Add nodes in $P_4$ into Open list\;
	}
}
\end{algorithm}

\end{document}