### 2024-01-25 Paper/experiments.tex

\section{Experimental Evaluation}

We implemented a prototype of {\toolname} in Python 3.8.  We conducted our evaluation on an AMD 5975WX  ($32$ cores$@3.6$GHz, 7nm, year 2022) with 256 GB of main memory and 2 NVIDIA RTX 3090. Gurobi 9.52 was used for solving MILP and LP problems.

We conducted our evaluation on neural networks trained on MNIST that have been employed in prior work. The networks have varying sizes: $5\times 100$, $5\times 200$, $8 \times 100$, and $8 \times 200$. It is worth remarking that despite seemingly small size, these instances are challenging for the current state of the art techniques. In particular, the current state of the art methods are unable to characterize $12\%$ to $20\%$ of the images as neither certified robust \cite{crown} nor find an adversarial example \cite{attack}. Following methodology in prior work \cite{prima,crown}, we evaluate on the first 1000 images of the dataset. We also experimented on MLP of size $6\times 500$ (results not reported in \cite{prima,crown}), checking the first 200 images of the dataset, given the computational overhead associated with such larger networks. Each DNN is associated with a $\varepsilon>0$ for the $L^\infty$ norm. 
%Each DNN is associated with a $\varepsilon>0$ for the $L^\infty$ norm.

\paragraph{Baseline}: 
We compare the following verifiers performing value-abstraction:
\begin{itemize}
	\item DeepPoly \cite{deeppoly}/ CROWN \cite{crown}. We report the runtime from our own implementation used in Compensate($K$). It shows that our purely Python implementation is not optimal ($10$ to $100$ times slower than the implementation in ($\alpha$)($\beta$)CROWN, ERAN and PRIMA). 
	\item PRIMA~\cite{prima} and $(\alpha)\beta$ CROWN. We use the results reported in \cite{prima} and \cite{crown} respectively for instances that were also used in prior works and report results for MLP $6\times500$ instances as it was not considered in prior work.

%	They use GPU while we do not. When computing the whole correlated space, 
%	parallelization is difficult.
%	Notice that on these DNNs, PRIMA resort to refined bounds on the first few layers (3 or 4) using exact MILP encoding of all the nodes (which is easily parallelized - 16 nodes considered in parallel in PRIMA). This is doable on the first few layers as the number of integer variable is not too large. By comparison, we run MILP on all the layers, but with a bounded number of integer variables to control the runtime and to scale to all the layers.
	\item ($\alpha$)$\beta$-CROWN, with results as reported in \cite{crown}, as it is the most accurate of the efficient tool to date. 
%	Its main idea is to implement a very efficient Branch and Bound (BaB) verification tool, subsuming BaBSR \cite{BaB}, and using GPU implementation (while we do not). Performance of pure BaB on these 
%	"hard to verify" DNNs is however impaired by the large number of branchs to consider. Similarly as PRIMA, $\beta$-CROWN resorts to refining the bounds of the first few layers using an exact MILP encoding, also computing 16 nodes in parallel, before running the branching heuristic BaB-FSB. 
%	\item We also experimented with the latest versions of $\alpha$-$\beta$-CROWN (October 2023) and PRIMA (2022) on MLP $6 \times 500$ (results not reported in \cite{crown,prima}).
	\item It is worth remarking that we do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI}, MIPplanet \cite{MIPplanet}, as $(\alpha)\beta$-CROWN has been shown to outperform these methods\cite{crown}.

\end{itemize}

The  objective of our evaluation was to answer the following  questions:

\begin{enumerate}
	\item How does the the choice of  the set  $Z$ impacts the accuracy of $\MILP_Z$? 
	\item  How does the performance, measured as the percentage of images verified, of {\toolname} compare to that of prior state of the art approaches? 
	\item Evaluate how the runtime of compensate($K$) scale with the size of DNNs.
\end{enumerate}

\subsection{Utility of Heuristic based on Compensation Strength  }

%The first group of experiments we run, tests the accuracy of choosing a set $Z$ of $K$ nodes using compensation strength, compared with randomly choosing the same number $K$ of nodes,  to understand if the choice is meaningful or any set $Z$ with the same size will result in a similar accuracy.
To measure the impact of the heuristic based on compensation strength, we focused on the smallest MLP (i.e. of size $5\times 100$ \cite{crown}) so as to obtain exact bounds for the first few layers using a full MILP encoding of the DNN. We test over the $\vx=59$th image in the MNIST dataset, as it has a large number of unstable ReLU nodes in the first few layers ($61$ in the first and $55$ in the second layer), so we can experiment with a larger choice of values) for $K$.

To measure the accuracy, we measure the uncertainty of all nodes in a layer:
the uncertainty of a node is the range between its computed lower and upper bound. 
We then average the uncertainty among all the nodes of the layer.
Formally, for a node $n$ with bounds $[\alpha,\beta]$, its uncertainty $unc(a) = \beta - \alpha$, and the average uncertainty of a layer $l$ is $\dfrac{\sum_{a\in l} unc(a)}{|l|}.$

\subsubsection*{ReLU Node in the first hidden layer}

\begin{table}[b!]
	\centering
	\begin{tabular}{|c||c|c|}
		\hline
		\text{Number $K$ of nodes in $Z$}  &  \text{Compensate strength} & \text{Random Choice}  \\ \hline
		\hline
		0  &  2.22 & 2.22  \\ \hline
		10  &  1.84 & 2.03  \\ \hline
		20  &  1.50 & 1.82  \\ \hline
		30  &  1.28 & 1.62  \\ \hline
		40  &  1.14 & 1.44  \\ \hline
		50  &  1.06 & 1.23  \\ \hline
		61 (max) & 1.05 &  1.05 \\ \hline
	\end{tabular}
	\caption{Average uncertainty of $\MILP_Z$ for nodes of the second layer, for $Z$ with $K$ ReLU nodes of the first layer (compensating strength vs random choice).}
	\label{tab:example0}
	%\vspace{-0.8cm}
\end{table}

%KSM: I don't think this is really the place for the paragraph below
% Let $c$ be a node of the second layer.
%A compensating pair $(\pi,\pi')$ of paths with target $c$ is thus of the form a node $a b c$ and $a b' c$, with $a$ an input node and $b,b'$ on the first layer. The compensating strength for $(b,b')$ is defined as $comp(b,b')=\sum_a \val_{\vx}(a) \min(weight(abc),weight(ab'c))$. Indeed, the pair $(a b c,a b' c)$ of path cannot compensate more than $\val_{\vx}(a) \min(weight(a,b,c),$\newline $weight(a,b',c))$, and $b,b'$ would help compensated the set of pairs of paths $\{(a b c,a b' c) \mid a$ is in the first layer$\}$. After selecting the heaviest $(b_0,b_0')$, we can then compute $comp(b)= \sum_{b' \text{already selected}} comp(b,b')+comp(b',b)$ and select iteratively the heaviest $b$'s one by one till reaching the threshold. 
%We do this for all node $c$ of the second layer.

We first focus on the choice of nodes in the first hidden layer, and its consequences on the accuracy of nodes in the second layer. 
We report in Table \ref{tab:example0} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly. The range of uncertainty created by inacurate computations is up to $1.17=2.22-1.05$, $2.22$ being the accuracy provided by LP, and $1.05$ by an exact MILP encoding of all the unstable ReLU nodes. Selecting $30$ out of the $61$ unstable ReLU nodes, the uncertainty created by inacurracy is reduced by $80\%$, while a random choice of nodes only leads up to roughly halving this number. We can deduce that selecting nodes based on compensating strength helps to select important ReLU nodes for the accuracy.

\subsubsection*{ReLU Node in the first and second hidden layer}

We now turn to evaluating the choice of ReLU nodes in two layers, focusing on the uncertainty of nodes in the third layers, wrt ReLU nodes in the first and second layer.
The bounds for nodes of the first two layers are computed exactly using the full MILP encoding. Let $d$ be a neuron of the third layer.
We keep the previous evaluation for ranking nodes in the previous (second) layer. 
Once the set $Z$ of selected nodes has sufficiently many nodes $c,c'$ in the second layer, adding $b,b'$ from the first layer to $Z$ allows to take into account accurately the pairs of paths of the form $(a b c d, a b' c' d)$ for all $\{c,c'\} \in Z$. We thus define:
\vspace{-0.2cm}
$$comp(b,b')=\sum_a \sum_{c,c' \in Z} \val_{\vx}(a) \min( weight(abcd),weight(ab'c'd))$$ 
\vspace{-0.2cm}

and the associated $comp(b)$, and select iteratively nodes $b$ (first layer) or $c$ (second layer) by selecting the heaviest $comp(b)$ or $comp(c)$.

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer $2$) only is less accurate than 
also choosing nodes in layer $1$. When picking in both layer $1$ and $2$, choosing $50$ out of $116$ unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inaccurate computations ($0.059$ vs $0.574$), while it only reduces the uncertainty by $65\%$ when choosing nodes randomly. Notice that applying this layer after layer will even further reduce uncertainty created by accumulation of inaccuracies. 
Last, we did not experiment for selecting ReLU nodes in 3+ layers earlier because the trade-off of accuracy vs evaluating the compensating strength of such paths is unclear.

\begin{table}[t!]	
	\centering
	\begin{tabular}{|c||c|c|c|}
		\hline
		\text{Number $K$}  &  \text{Compensate layer} 1+2 &  \text{Compensate layer} 2 & \text{Random layer } 1+2 \\ \hline
		\hline
		0  &  1.758 & 1.758 & 1.758  \\ \hline
		10  &  1.677 & 1.677 & 1.695  \\ \hline
		20  &  1.539 & 1.567 & 1.622  \\ \hline
		30  &  1.431 & 1.489 & 1.542  \\ \hline
		40  &  1.329 & 1.454 & 1.47  \\ \hline
		50  &  1.243 & 1.432 & 1.39  \\ \hline
		116 (max) &  1.184 & 1.422 & 1.184  \\ \hline
	\end{tabular}
	\caption{Average uncertainty of $\MILP_Z$ for nodes of the third layer, for $Z$ with $K$ ReLU nodes of the 1st and 2nd layer (compensating strength vs random choice).}
	\label{tab:example1}
	\vspace{-0.8cm}
\end{table}

\subsection{Comparison with value-abstraction Verifiers}

In Table \ref{tab:example}, we report the accuracy and runtime of these different methods.
DeepPoly/CROWN is by far the fastest, but also the most inaccurate, certifying robustness for less than $30\%$ of the images, whereas there are at least $82\%$ of the images for which no attacked is found (attacks computed by \cite{attack} in $\beta$-CROWN).
This is because these DNNs, although of reasonable size, are hard to verify (they are trained in a natural way, with high compensation strength, see Section \ref{Sec.comp}).

Compared with PRIMA (using refinement of the first few layers by exact MILP), the overall pipeline is quite similar, looking at all the nodes one by one from the beginning of the network till the end. Our method compare favorably, both in terms of time and accuracy: our fast adaptive setting is always faster than PRIMA, sometimes by a large margin ($>3$ times faster on $8 \times 100$ while verifying $10\%$ more images), and also more accurate (except for $5 \times 200$ which is quite pointless on this shallow DNN - the fast fixed method is both faster and more accurate than PRIMA), sometimes by a large margin (by $14 \%$  for $5 \times 100$ while being $30\%$ faster). Further, unlike PRIMA, we do not resort to a GPU, and our purely Python implementation is not optimal. This is very promising for this idea of compensation, which indirectly treats dependencies between the variables, which PRIMA represents explicitly.

\begin{table}[t!]
	\centering
	\begin{tabular}{|c||c|c|c||cc|cc||c|}

		\hline
		\text{DNN}  & DeepPoly & PRIMA & $\beta$-CROWN &\multicolumn{2}{@{}c@{}|}{\color{blue}\text{ Comp (adapt.) }} & \multicolumn{2}{@{}c@{}|}{\color{blue} \text{ Comp (fixed) }} & Upper\\ 
		& / CROWN & (refined) & (refined BaB) & {\color{blue}fast} & {\color{blue}slow} & {\color{blue}fast} &{\color{blue}slow} & Bound\\
		\hline \hline

		$5 \times100$\  &   $16\%$ & $51\%$ & $\mathbf{69.9\%}$ & {\color{blue}$57.5\%$} & {\color{blue}$66.1\%$} & {\color{blue}$65.3\%$} & {\color{blue}$68.4\%$} &  $84.2 \%$ \\ 
		$\epsilon = 0.026$ & 5s & 159s & 102s& {\color{blue}75s} & {\color{blue}140s} & {\color{blue}117s} & {\color{blue}142s} &  \\
		\hline

		$5 \times 200$ \  &  $18.2\%$ & $69\%$ & $\mathbf{77.4\%}$ & {\color{blue}$64.3\%$} & {\color{blue}$74.7\%$} & {\color{blue}$72\%$} & {\color{blue}$76.8\%$} & 
		$90.1 \%$\\ 
		$\epsilon = 0.015$ & 11s & 224s & 86s & {\color{blue}163s} & {\color{blue}285s} & {\color{blue}206s} & {\color{blue}279s}  &\\ \hline \hline

		$8\times100$\  &   $29.2\%$ & $42.8\%$ & $62\%$ & {\color{blue}$52.7\%$} & {\color{blue}$59.7\%$} & {\color{blue}$61.4\%$} & {\color{blue}$\mathbf{65.8\%}$}  & $82.0 \%$ \\ 
		$\epsilon = 0.026$ & 8s & 301s & 103s & {\color{blue}92.3s} & {\color{blue}179s} & {\color{blue}170s} & {\color{blue}346s} & \\
		\hline

		$8\times200$\  &   $25.9\%$ & $62.4\%$ & $73.5\%$ & {\color{blue}$68.1\%$} & 
		{\color{blue}$79.1\%$} & {\color{blue}$72.6\%$} & {\color{blue}$\mathbf{80.7\%}$} & $91.1 \%$ \\ 
		$\epsilon = 0.015$ & 17s & 395s & 95s  & {\color{blue}297s} & {\color{blue}535s} & {\color{blue}$480s$} & {\color{blue}697s$^\star$} & \\ \hline

		%6$\times$500\  &   0.035& 1.758 & time & 1.758 & time &    1.758 & time & 1.758 & time & \\ 
		%$\epsilon = 0.035$ & & & & & & & & & &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 1000 images. 
		Results for PRIMA, $\beta$-Crown and the upper bound on the $\%$ are from \cite{crown}.
		\newline $^\star$ For $8 \times 200$, slow fixed results are obtained after running slow adaptative.}
	\label{tab:example}
	\vspace{-1cm}
\end{table}

Compared with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), the picture is more balanced, because the fundamentals of both algorithms are different. BaB focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence Compensate($K$) will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized $\beta$-CROWN: our slow setting is $1.4$ to 
$6.5$ times slower than $\beta$-CROWN. Notice Compensate($K$) it is not using GPUs. 
Concerning accuracy: on shallower DNNs ($5 \times 100$ and $5 \times 200$ with $5$ hidden layers), refined BaB uses MILP on all but the last $2$ layers, leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accurate than using the slow fixed setting of Compensate($K$), with a small difference of less than $1.5\%$. On deeper DNNs ($8 \times 100$ and $8 \times 200$ with $8$ hidden layers), the full MILP used in refined BaB cannot treat the last $4$ layers, and BaB has more ReLU nodes to branch on: 
the most accurate setting (slow fixed) of Compensate($K$) is more accurate than $\beta$-CROWN, closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $10.4\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than in Compensate($K$) (see Section \ref{Discussion}).

Importantly, ($\alpha$)$\beta$-Crown, while efficient, can face a (complexity) wall, when the number of branches becomes too extreme for BaB to work: either BaB succeds fast, or it will not suceed even with very long timeouts ({\em refined} BaB is more accurate than pureBaB on these networks). We verify that in Table \ref{tab:example3}, on a larger DNN learnt in a natural way, namely MLP $6 \times 500$.
We tested PureBaB with two timeouts (TO), from standard 30s to an extreme 2000s, to match the average runtime of our adaptative implementation (fast mode). Only $3.5\%$ more images could be ceritifed by extending the TO from 30s to 2000s, certifying $20\%$ less images than our prototype in the same time. Notice that the refined BaB mode of ($\alpha$)$\beta$-Crown (first few layers using MILP) failed to run on this DNN (error code: NoImplementation), and we were unable to fix the problem, which is not really surprising as running exact MILP with so many integer variables is not likely to be efficient, as confirmed while running PRIMA (refined).

% test

\begin{table}[t!]
	\centering
	\begin{tabular}{|c||c|c|cc||c||c|}

		\hline
		\text{DNN}  & DeepPoly & PRIMA &  \multicolumn{2}{@{}c@{}|}{\text{ $\alpha$-$\beta$-CROWN (pureBaB) \,}} & 
		{\color{blue}Comp (adapt.)} & Upper\\ 
		& / CROWN & (refined) & TO=30s & TO=2000s & {\color{blue}fast} & Bound\\
		\hline \hline
		$6\times500$\  &   $26\%$ & $32.5\%$ & $41\%$ & $44.5\%$  & {\color{blue}$\mathbf{65\%}$} & $92\%$ \\ 
		$\epsilon = 0.035$ & 34s & 119s & 18.4s & 954s & {\color{blue}925s} &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}

\end{table}

\begin{figure}[h!]
	\centering
	\begin{tikzpicture}
		\begin{axis}[width=10cm,height=5cm,
			xlabel={},
			ylabel={},
			legend pos=south east,
			ymin=0,
			xtick={1,2,3,4,5}, 
			xticklabels={$5\times 100$, $8\times 100$, $5\times 200$, $8\times 200$, $6\times 500$},
			]

			\addplot[mark=triangle, blue] coordinates {
				(1, 0.234)
				(2, 0.2125)
				(3, 0.206)
				(4, 0.3)
			};
			\addlegendentry{fast fixed}

			\addplot[mark=square, red] coordinates {
				(1, 0.15)
				(2, 0.115)
				(3, 0.163)
				(4, 0.1856)
				(5, 0.308)
			};
			\addlegendentry{fast adaptative}

		\end{axis}
	\end{tikzpicture}
	\caption{runtime (seconds per node)}
	\label{fig2}
\end{figure}

Last, we display the average runtime per number of nodes of both fast adaptive and fast fixed compensation, for the different DNNs considered. The runtime per node is not constant (due to $\LP(N,K)$), but it is still controlled and does not show an exponential blow-up, from 0.1s per node for the smallest DNN to 0.3s per node for the biggest DNN. This indicates that Compensate($K$) can be run on larger DNNs without hitting the same complexity wall as pure BaB.

\begin{table}[]
	\centering
	\begin{tabular}{|c|c|cc||cc||c|}
		\hline
		\text{DNN}  & DeepPoly &  \multicolumn{2}{@{}c@{}|}{\text{ $\alpha$-$\beta$-CROWN (refinedBaB) \,}} & 
		\multicolumn{2}{@{}c@{}|}{\text{ \color{blue}\CMP slow \,}}
		& Upper\\ 
		& / CROWN & TO=300s & TO=3000s & {\color{blue}adapt} & {\color{blue}fixed} & Bound\\
		\hline \hline
		$8\times200$\  &   $25\%$? & $72\%$? & $\mathbf{79.5\%}$ & {\color{blue}$75.5\%$}  & {\color{blue}$77\%$} & $91\%$ \\ 
		$\epsilon = 0.015$ & 17s & 95s? & 451s & {\color{blue}391s} & {\color{blue}562s} &\\ \hline
	\end{tabular}
	\caption{WIP! $\%$ of verified images over 200 images. To see how refined BAB scale with TimeOut.}
	\label{tab:example4}
\end{table}

\section{Limitations and Discussion}
\label{Discussion}

For DNNs robustly trained, $\alpha$-$\beta$-CROWN and PRIMA are excellent solutions, both very fast and very accurate, usually matching or getting very close \cite{crown} to the upper bound on the $\%$ of verified images \cite{attack}. Compensation strength is not required in these contexts, reason why we did not consider it.

We only considered DNNs with ReLU activation functions in this paper. The reason is that MILP, which is the basic block of Compensate($K$) (and BaB for $\beta$-Crown) can handle the ReLU activation function efficiently. Other activation functions are abstracted in more complex ways, e.g. by PRIMA (unlike $\beta$-CROWN). Still, the compensation paradigm is  orthogonal to the activation function. We believe that studying the compensation strength would be useful with other activation functions as well. We leave that for future work.

Accuracy of Compensate($K$) is among the highest of the efficient verifiers that do abstraction on values. Runtime, while never blowing-up, can be on the slow side though, in particular for larger DNNs. There are many optimizations possible. The simplest is to run $\alpha$-$\beta$-CROWN first in the pipeline, and only use Compensate($K$) for those images for which $\alpha$-$\beta$-CROWN cannot conclude, possibly using more CPU cores to compute more nodes in parallel. Another simple way is to modify the {\em refined} BaB of $\beta$-CROWN to run Compensate($K$) on more layers than what exact MILP can do, and finish the verification using BaB. More involved would be to modify BaB to branch on unstable ReLU nodes guided by the compensation strength. Also, an analysis of important nodes for the uncertain output nodes could be performed, allowing to compute much faster less accurate bounds for those nodes not important. Last, one could perform a Refinement Abstraction framework similar to \cite{atva}, \cite{elboher}, or \cite{SRGR}.

On the correlation between neurons, which PRIMA encodes explicitly, our results tend to show that it is more efficient to use the compensation strength to abstract the network. Still, considering ReLU nodes exactly tends to be very accurate locally, but less when the source of correlation is many layers away. Few important correlations {\em a la} PRIMA could be kept explicitly as linear constraints in the MILP encoding, helping to reduce the inaccuracies further.