### 2024-02-16 Reports/tech_report1.tex
Tech Report draft

\documentclass{llncs}
\pagestyle{plain}

%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\vW}{\boldsymbol{W}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vb}{\boldsymbol{bias}}
\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}

\newcommand{\UB}{\mathrm{UB}}
\newcommand{\LB}{\mathrm{LB}}
\newcommand{\ub}{\mathrm{ub}}
\newcommand{\lb}{\mathrm{lb}}
\newcommand{\B}{\mathrm{B}}

\newcommand{\CMP}{{\textrm{CMP}}\ }

\newcommand{\toolname}{\ensuremath{\mathsf{Comp}}}

%\usepackage{amsmath, amsthm, amssymb, amsfonts}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{corollary}{Corollary}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}

\newcommand{\ReLU}{\mathrm{ReLU}}

\title{Technical Report}
\date{}

\begin{document}

\maketitle

\section*{Introduction}

Only formal experiments will be recorded. Pre-debugging will not be recorded unless necessary.

\section{February 15 experiments}

\subsection{Rerun the gamma test}

This experiment is to rerun the interrupted experiment on14 February night.

\vspace*{1ex}

The running code's file name is \verb*|DiamondLipschitz_all_layers_test_LP_use_bound.py|. And the output file name is \verb*|Gamma_V0p2_all_layers_MILP_usebounds1.txt|. It is to compute the gamma ($\gamma$). It is based on the new hyperplane constraints, use MILP model to compute the gamma from layer 5 to layer 5.

\subsubsection*{Parameters.}

\begin{itemize}
	\item \emph{Experiments Id}: 021500102024

	\item\emph{Network:} 6x100 MNIST normal.

	\item\emph{Image:} 144 (meaningless).

	\item\emph{Epsilon $\varepsilon$:} 0.001.

	\item\emph{Timeout:} No global timeout. The loop timeout for $\gamma$ is 600s with 12 loops at most.

	\item\emph{Open $\ReLU$ number:} 17 for layer 3, 23 for layer 5, 35 for layer 7 and 9.

	\item\emph{Processes:} 20 processes in parallel.

	\item\emph{Read binary files:} \verb*|dict_bounds_gurobi_alphabeta_2024021401.b|; this file stores the alpha beta of nodes of all layers for the same network. \verb*|dict_bounds_gurobi_gamma_2024021501.b|; this file stores the gamma of nodes up to layer 3, computed by the same code.
\end{itemize}

\subsubsection*{Technical Sketch.}

The new group of hyperplane constraints is as follows: \begin{align*}
	&\hat{x}_i - \hat{x}_i' \leq \frac{1}{2} (x_i-x_i') + \frac{1}{2} \gamma_i\\
	&\hat{x}_i' - \hat{x}_i \leq \frac{1}{2} (x_i'-x_i) + \frac{1}{2}\gamma_i
\end{align*} Here, $x_i$ and $x_i'$ are the two variables corresponding to a node $i$ in previous layer, and $\gamma_i$ is the $\gamma$ of node $i$ computed in previous layer; $\hat{x}_i,\hat{x}_i'$ are variables after $\ReLU$ function.

One characteristic is that the constraints do not reply on $\alpha$ or $\beta$. So even with $\alpha,\beta$ of very poor quality, this group of constraints can work.

This running recover the last layer: in early experiments, the last layer is modified to 9 nodes. Now it is recovered to 10 nodes. Because of this, the Bootstrap data of last layer is computed.

\subsubsection*{Results.}

No result.

It seems that open-23-node for layer 5 with new hyperplane constraints is not good, at least with respect to the time cost: most nodes run out of 600s x 12, and this means we need 10 hours to do the whole layer 5.

Let's try a fewer number of open $\ReLU$ nodes.

\subsection{Rerun the gamma test: layer 5 with changed parameters}

This experiments shares some setting as previous one.

\subsubsection*{Parameters.}

\begin{itemize}
	\item\emph{Experiments Id}: 021500202024

\item\emph{Network:} 6x100 MNIST normal.

\item\emph{Image:} 144 (meaningless).

\item\emph{Epsilon $\varepsilon$:} 0.001.

\item\emph{Timeout:} No global timeout. The loop timeout for $\gamma$ is 600s with 12 loops at most.

\item\emph{Open $\ReLU$ number:} 17 for layer 5.

\item\emph{Processes:} 20 processes in parallel.

\item\emph{Read binary files:} \verb*|dict_bounds_gurobi_alphabeta_2024021401.b|; this file stores the alpha beta of nodes of all layers for the same network. \verb*|dict_bounds_gurobi_gamma_2024021501.b|; this file stores the gamma of nodes up to layer 3, computed by the same code.
\end{itemize}

\subsubsection*{Technical Sketch.}

The same as previous experiments.

\subsubsection*{Results.}

\begin{itemize}
	\item \emph{Log:} \verb*|Gamma_V0p2_all_layers_MILP_usebounds1.txt|

	\item \emph{Total time:} 4779s.

	\item \emph{Key results:} $\gamma$ for layer 5 is 0.9922133701854773.
\end{itemize}

Open 17 nodes is OK, but we can still increase the number of open ReLU nodes for layer 5. Perhaps open 20 nodes is a good choice.

\subsection{Rerun the gamma test: layer 7,9 with changed parameters}

This experiments shares some setting as previous ones.

\subsubsection*{Parameters.}

\begin{itemize}
	\item\emph{Experiments Id}: 021500302024

	\item\emph{Network:} 6x100 MNIST normal.

	\item\emph{Image:} 144 (meaningless).

	\item\emph{Epsilon $\varepsilon$:} 0.001.

	\item\emph{Timeout:} No global timeout. The loop timeout for $\gamma$ is 600s with 12 loops at most.

	\item\emph{Open $\ReLU$ number:} 20 for layer 7 and 9.

	\item\emph{Processes:} 20 processes in parallel.

	\item\emph{Read binary files:} \verb*|Diamond_record_6_100_dictionary_with_end_layer5_2024_02_15_11_40_01.b|; this file stores all alpha beta gamma of nodes of all layers for the same network (gamma is up to layer 5). 
\end{itemize}

\subsubsection*{Technical Sketch.}

The same as Experiments 021500202024.

\subsubsection*{Results.}

\begin{itemize}
	\item \emph{Log:} \verb*|Gamma_V0p2_all_layers_MILP_usebounds2.txt|

	\item \emph{Total time:} 14985.4s.

	\item \emph{Key results:} $\gamma$ for layer 7 is 4.177719108178739; $\gamma$ for layer 9 is 16.878953032158446.
\end{itemize}

\subsection{Gamma test for robust 5x100 using MILP}

This experiment is to rerun the interrupted experiment on14 February night.

\vspace*{1ex}

The running code's file name is \verb*|DiamondLipschitz_all_layers_test_LP_DiffAI.py|. The output file name is \verb*|Gamma_V0p2_all_layers_5x100_robust_MILP.txt| and \verb*|Diamond_record_5_100_dictionary_with_end_layer9_2024_02_15_19_29_17.b|. It is to compute the gamma ($\gamma$). It is based on the new hyperplane constraints, use MILP model to compute the gamma of robust 5x100 from layer 1 to layer 9.

\subsubsection*{Parameters.}

\begin{itemize}
	\item \emph{Experiments Id}: 021500402024

	\item\emph{Network:} 6x100 MNIST robust.

	\item\emph{Image:} 144 (meaningless).

	\item\emph{Epsilon $\varepsilon$:} 0.001.

	\item\emph{Timeout:} No global timeout. The loop timeout for $\gamma$ is 600s with 12 loops at most.

	\item\emph{Open $\ReLU$ number:} 17 for layer 3, 23 for layer 5, 35 for layer 7 and 9.

	\item\emph{Processes:} 20 processes in parallel.

	\item\emph{Read binary files:} No read file. I planned to read alpha-beta data, but the pre-computed alpha-beta data seems to have some problems, so I use the version that will compute the alpha beta during the running.
\end{itemize}

\subsubsection*{Technical Sketch.}

The same as other experiments today.

\subsubsection*{Results.}

\begin{itemize}
	\item \emph{Log:} \verb*|Gamma_V0p2_all_layers_5x100_robust_MILP.txt|

	\item \emph{Total time:} 19270.8s.

	\item \emph{Key results:} $\gamma$ for layer 1 to layer 9: 0.006884580440378905, 0.014342863034246162,
	0.034691083484486775, 0.09665049332723069,
	0.28105516604910674
\end{itemize}

\section{February 16 experiments}

\subsection{Rerun the gamma test: layer 3 with new settings}

\subsubsection*{Parameters.}

\begin{itemize}
	\item\emph{Experiments Id}: 021600102024

	\item\emph{Network:} 6x100 MNIST normal.

	\item\emph{Image:} 144 (meaningless).

	\item\emph{Epsilon $\varepsilon$:} 0.001.

	\item\emph{Timeout:} No global timeout. The loop timeout for $\gamma$ is 600s with 12 loops at most.

	\item\emph{Open $\ReLU$ number:}

	\item\emph{Processes:} 20 processes in parallel.

	\item\emph{Read binary files:} 
\end{itemize}

\subsubsection*{Technical Sketch.}

The same as Feb 15's experiments.

\subsubsection*{Results.}

\end{document}