### 2025-01-25 CAV 2025 TODO/Comparison.tex

\section{Comparison of other methods.}

\label{sec4p5}

Here we list the FSB formula from the paper \cite{FSB}:

\begin{align*}
	\bm{\bar{\lambda}}_{n-1} =& -W^T_n\bm{1}, \hspace*{10ex}  	\bm{\bar{\lambda}}_{k-1} = W^T_k\big(\dfrac{[\hat{\bm{u}_k}]_+}{[\hat{\bm{u}_k}]_+-[\hat{\bm{l}_k}]_-}\odot\bm{\bar{\lambda}}_{k}\big) \hspace*{-10ex}  &k\in [2,n-1]\\
	\bm{s}_{SR} =& \lvert\max\{0,\bm{\bar{\lambda}}_{k}\odot\bm{b}_{k}\}-\frac{\hat{\bm{u}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}}_k\odot\bm{\bar{\lambda}}_{k}\odot\bm{{b}}_{k}+\frac{\hat{\bm{u}}_k\odot\hat{\bm{l}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}_k}\odot[\bm{\bar{\lambda}}_{k}]_+\rvert &\\
	&\odot\bm{1}_{\hat{\bm{u}}_k>0,\hat{\bm{l}}_k<0}&k\in [1,n-1]\\
	\bm{s}_{FSB} =& \lvert\min\{0,\bm{\bar{\lambda}}_{k}\odot\bm{b}_{k}\}-\frac{\hat{\bm{u}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}}_k\odot\bm{\bar{\lambda}}_{k}\odot\bm{{b}}_{k}+\frac{\hat{\bm{u}}_k\odot\hat{\bm{l}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}_k}\odot[\bm{\bar{\lambda}}_{k}]_+\rvert &\\
	&\odot\bm{1}_{\hat{\bm{u}}_k>0,\hat{\bm{l}}_k<0}&k\in [1,n-1]\\
\end{align*}

\begin{figure}[t!]
	\centering
	\begin{tikzpicture}

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (input1) {$n_1$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (input2) at ($(input1) + (0,-1.5)$) {$n_2$};

		% Hidden layers

		\node (hidden10) at ($(input1) + (2.5,0.6)$) {$0$};

		\node (hidden20) at ($(input1) + (2.5,-1.5-0.6)$) {$0$};

		\node (hidden50) at ($(input1) + (7.5,0.6)$) {$0$};

		\node (hidden60) at ($(input1) + (7.5,-1.5-0.6)$) {$1.5$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (hidden1) at ($(input1) + (2.5,0)$) {$n_3$};
		\node[circle, draw= black, thick] (hidden2) at ($(input1) + (2.5,-1.5)$) {$n_4$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (hidden3) at ($(input1) + (5,0)$){$\hat{n}_3$};
		\node[circle, draw= black, thick] (hidden4) at ($(input1) + (5,-1.5)$) {$\hat{n}_4$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (hidden5) at ($(input1) + (7.5,0)$){$n_5$};
		\node[circle, draw= black, thick] (hidden6) at ($(input1) + (7.5,-1.5)$) {$n_6$};

		% Output layer
		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (output1) at ($(input1) + (10,0)$){$\hat{n}_5$};

		\node[circle, draw= black, thick, minimum width = 20,
		minimum height = 20] (output2) at ($(input1) + (10,-1.5)$){$\hat{n}_{6}$};

		% Connections

		\draw[->,thick] ($(input1) + (-1.5,0)$) -- (input1) node[midway, above] {$[-1,1]$};

		\draw[->,thick] ($(input1) + (-1.5,-1.5)$) -- (input2) node[midway, above] {$[-1,1]$};

		\draw[->,thick] (input1) -- (hidden1) node[near start, above] {$1$};
		\draw[->,thick] (input1) -- (hidden2)node[near start, above] {$1$};

		\draw[->,thick] (input2) -- (hidden1) node[near start, below] {$1$};
		\draw[->, thick] (input2) -- (hidden2)node[near start, below] {$-1$};

		\draw[->, thick] (hidden1) -- (hidden3) node[midway, above] {$\max(n_1,0)$};
		\draw[->, thick] (hidden2) -- (hidden4) node[midway, above] {$\max(n_2,0)$};

		\draw[->, thick] (hidden3) -- (hidden5) node[near start, above] {$1$};			
		\draw[->,thick] (hidden3) -- (hidden6) node[near start, above] {$0$};

		\draw[->,thick] (hidden4) -- (hidden5)node[near start, below] {$2$};
		\draw[->,thick] (hidden4) -- (hidden6)node[near start, below] {$1$};

		\draw[->,thick] (hidden5) -- (output1) node[midway, above] {$\max(n_5,0)$};
		\draw[->,thick] (hidden6) -- (output2) node[midway, above] {$\max(n_6,0)$};

	\end{tikzpicture}
	\caption{}
\end{figure}

\begin{tikzpicture}[scale=1, >=stealth]

	% Draw axes
	\draw[->] (-5,0) -- (4,0) node[right] {$x$};
	\draw[->] (0,-1) -- (0,3) node[above] {$y$};

	% Draw ReLU function
	\draw[line width=0.4mm, blue] (-3,0) -- (0,0);
	\draw[thick, blue] (0,0) -- (2.5,2.5) node[below, shift={(0.5,-0.4)}] {$y = \ReLU(x)$};
	\draw[thick, blue] (-3,0) -- (2.5,2.5) node[above, shift={(-0.5,0.2)}] {$y = \frac{\UB}{\UB-\LB} x-\frac{\UB\LB}{\UB-\LB}$};

	% Add labels
	\draw[dashed] (2.5,0) -- (2.5,2.5) -- (0,2.5); % Optional grid
	\node[below left] at (0,0) {$0$};

	% Add tick marks

	\foreach \x in {2.5}
	\draw[shift={(\x,0)}] (0,0.1) -- (0,-0.1) node[below] {$\UB$};
	\foreach \x in {-3}
	\draw[shift={(\x,0)}] (0,0.1) -- (0,-0.1) node[below] {$\LB$};

	\foreach \y in {2.5}
	\draw[shift={(0,\y)}] (0.1,0) -- (-0.1,0) node[left] {$\UB$};

\end{tikzpicture}