### 2025-01-26 CAV 2025 TODO/Paper_noProof.tex
title abstract intro

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{url}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\vW}{\boldsymbol{W}}

\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}
\newcommand{\Improve}{\mathrm{Improve}}
\newcommand{\Utility}{\mathrm{Utility}}
\newcommand{\Sol}{\mathrm{Sol}}
\newcommand{\sol}{\mathrm{sol}}

\newcommand{\UB}{\mathrm{UB}}
\newcommand{\LB}{\mathrm{LB}}
\newcommand{\ub}{\mathrm{ub}}
\newcommand{\lb}{\mathrm{lb}}
\newcommand{\B}{\mathrm{B}}
\usepackage{amsmath, amssymb, amsfonts}

\newcommand{\ReLU}{\mathrm{ReLU}}

\newcommand{\CMP}{{\textrm{CMP}}\ }

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\toolname}{Hybrid MILP}

\title{Solution-aware vs global ReLU selection: \\
partial MILP strikes back for DNN verification}
\date{}

%\author{Antiquus S.~Hippocampus, Natalia Cerebro \& Amelie P. Amygdale \thanks{ Use footnote for providing further information
%		about author (webpage, alternative address)---\emph{not} for acknowledging
%		funding agencies.  Funding acknowledgements go at the end of the paper.} \\
%	Department of Computer Science\\
%	Cranberry-Lemon University\\
%	Pittsburgh, PA 15213, USA \\
%	\texttt{\{hippo,brain,jen\}@cs.cranberry-lemon.edu} \\
%	\And
%	Ji Q. Ren \& Yevgeny LeNet \\
%	Department of Computational Neuroscience \\
%	University of the Witwatersrand \\
%	Joburg, South Africa \\
%	\texttt{\{robot,net\}@wits.ac.za} \\
%	\AND
%	Coauthor \\
%	Affiliation \\
%	Address \\
%	\texttt{email}
%}

\begin{document}

	\maketitle

	\begin{abstract}
		Branch and Bound (BaB) is considered as the most efficient technique for DNN verification: it can propagate bounds over numerous branches, to accurately under- and over-approximate values a given neuron can take even in large DNNs, enabling formal verification of properties such as local robustness. Nevertheless, the number of branches grows {\em exponentially} with important variables, and there are complex instances for which the number of branches is too large to handle even using BaB. In these cases, providing more time to BaB is not efficient, as the number of branches treated is {\em linear} with the time-out. Such cases arise with verification-agnostic DNNs, non-local properties (e.g. global robustness, computing Lipschitz bound, compressed sensing), etc. 
		%The fact that pure BaB is not that efficient for e.g. verification-agnostic (even very small) DNNs has been witnessed before. The workaround, e.g. in {\em refined} $\alpha,\beta$-CROWN, was to precompute very accurate bounds for the first few neurons of the DNN using a complete full MILP encoding. Non-surprisingly, this very slow technique does not scale but to small DNNs.
		%Indeed, one of its implementation, $\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best trade-off between accuracy and runtime. 
		%VNNcomp however is focusing on relatively easy verification problems.

        To handle complex instances, we revisit a divide-and-conquer approach to break down the complexity: instead of few complex BaB calls, we rely on many small {\em partial} MILP calls. The crucial step is to select very few but very important ReLUs to treat using (costly) binary variables. The previous attempts were suboptimal in that respect. In this paper, we propose a novel {\em solution-aware} ReLU ranking (SAR), as well as adapt the BaB-SR and BaB-FSB branching functions as {\em global} ReLU ranking (GR) functions. We compare them theoretically as well as experimentally. Surprisingly perhaps, the most accurate solution (SAR) for {\em selecting} ReLUs to treat as binary variables is different from the most efficient solution (GR) to {\em rank} within this selection. Compared with previous attempts, SAR reduces the number of integer variables by around 4 times while maintaining the same level of accuracy. Implemented in {\em Hybrid MILP}, calling first $\alpha,\beta$-Crown with a short time-out to solve easier instances, and then partial MILP, produces a very accurate yet efficient verifier, reducing by up to $40\%$ the number of undecided instances to low levels ($8-15\%$), while keeping a reasonable runtime ($46s-417s$ on average per instance), even for fairly large CIFAR CNNs with 20 000 nodes.

		%a novel Utility function that selects few neurons to be encoded with accurate but costly integer variables in a {\em partial MILP} problem. The novelty resides in the use of the solution of {\em one} (efficient LP) solver to accurately compute a selection $\varepsilon$-optimal for a given input.
		%This allows us to carefully craft a {\em partial MILP} solution which selects automatically few neurons encoded as integer variables, the rest using the LP relaxation.

	\end{abstract}

\section{Introduction}

\input{introduction}
%
\input{relatedwork}

\section{Notations and Preliminaries}

In this paper, we will use lower case latin $a$ for scalars, bold $\boldsymbol{z}$ for vectors, 
capitalized bold $\boldsymbol{W}$ for matrices, similar to notations in \textcolor{blue}{\cite{crown}}.
To simplify the notations, we restrict the presentation to feed-forward, 
fully connected ReLU Deep Neural Networks (DNN for short), where the ReLU function is $ReLU : \mathbb{R} \rightarrow \mathbb{R}$ with
$ReLU(x)=x$ for $x \geq 0$ and $ReLU(x)=0$ for $x \leq 0$, which we extend componentwise on vectors.

%In this paper, we will not use tensors with a dimension higher than matrices: those will be flattened.

%\subsection{Neural Network and Verification}

% testtesttesttest
An $\ell$-layer DNN is provided by $\ell$ weight matrices 
$\boldsymbol{W}^i \in \mathbb{R}^{d_i\times d_{i-1}}$
and $\ell$ bias vectors $\vb^i \in \mathbb{R}^{d_i}$, for $i=1, \ldots, \ell$.
We call $d_i$ the number of neurons of hidden layer $i \in \{1, \ldots, \ell-1\}$,
$d_0$ the input dimension, and $d_\ell$ the output dimension.

Given an input vector $\boldsymbol{z}^0 \in \mathbb{R}^{d_0}$, 
denoting $\hat{\boldsymbol{z}}^{0}={\boldsymbol{z}}^0$, we define inductively the value vectors $\boldsymbol{z}^i,\hat{\vz}^i$ at layer $1 \leq i \leq \ell$ with
\begin{align*}
	\boldsymbol{z}^{i} = \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{i-1}+ \vb^i \qquad \, \qquad
	\hat{\boldsymbol{z}}^{i} = ReLU({\boldsymbol{z}}^i).
\end{align*}

The vector $\hat{\boldsymbol{z}}$ is called post-activation values, 
$\boldsymbol{z}$ is called pre-activation values, 
and $\boldsymbol{z}^{i}_j$ is used to call the $j$-th neuron in the $i$-th layer. 
For $\boldsymbol{x}=\vz^0$ the (vector of) input, we denote by $f(\boldsymbol{x})=\vz^\ell$ the output. Finally, pre- and post-activation neurons are called \emph{nodes}, and when we refer to a specific node/neuron, we use $a,b,c,d,n$ to denote them, and $W_{a,b} \in \mathbb{R}$ to denote the weight from neuron $a$ to $b$. Similarly, for input $\boldsymbol{x}$, we denote by $\val_{\boldsymbol{x}}(a)$ the value of neuron $a$ when the input is $\boldsymbol{x}$. A path $\pi$ is a sequence $\pi=(a_i)_{k \leq  i \leq k'}$ of neurons in consecutive layers, and the weight of $\pi$ is 
$weight(\pi)=W_{a_k,a_{k+1}} \times \cdots \times  W_{a_{k'-1},a_{k'}}$.

\iffalse
and the $i$-th hidden layer is a vector in $\mathbb{R}^{d_i}$, 
and the output layer is a vector in $\mathbb{R}^{d'}$ or a scale. 
The weights, bias and activation functions decide propagate the from previous to the next layer. In formula, from layer $l_{i-1}$ to layer $l_{i}$, the weight 
$\boldsymbol{W}^i$ is matrix of $d_i\times d_{i-1}$, 
the bias is a vector $\vb^i$ in $\mathbb{R}^{d_i}$, and the activation function 
is $\sigma$, then  if the $i-1$-th layer is $\hat{\boldsymbol{z}}^{(i-1)}$, 
then the value of $i$-th layer is computed by: 
\begin{align*}
	{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \vb^i\\
	\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)).
\end{align*} The vector $\hat{\boldsymbol{z}}$ is called post-activation values, and $\boldsymbol{z}$ is called pre-activation values, and $\boldsymbol{z}^{(i)}_j$ is used to call the $j$-th neuron in the $i$-th layer. In our style, we also call neurons \emph{nodes} and use $a,b,c,d$ to denote them. We use $W_{ab}$ to denote the weight from neuron $b$ to $a$. We use $\boldsymbol{x}$ to denote the vector of input and  $f(\boldsymbol{x})$ to denote the output.
\fi

\medskip

Concerning the verification problem, we focus on the well studied local-robustness question. Local robustness asks to determine whether the output of a neural network will be affected under small perturbations to the input. 
Formally, for an input $\vx$ perturbed by $\varepsilon >0$ under distance $d$, then the DNN is locally $\varepsilon$-robust in $\vx$ whenever:
\begin{align*}
	\forall \boldsymbol{x'} \text{ s.t. } d(\vx,\vx')\leq \varepsilon, \text{ we have }  
	argmax_i (f(\boldsymbol{x'})[i]) = argmax_i(f(\boldsymbol{x})[i])
\end{align*}

\iffalse
In some cases, the output is a vector but the aim to get the label of dimension with the minimal value. In this case, the problem can be written as:\begin{align*}
	\forall \boldsymbol{x} \in\mathcal{D} \  \min f(\boldsymbol{x}) = \min f(\boldsymbol{x}_0)
\end{align*}

If so, the question of verification can turn to the following optimization question: \begin{align*}
	\min f(\boldsymbol{x}) \ s.t. {\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ b^i\\
	\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)), \boldsymbol{x}\in\mathcal{D}.
\end{align*}

In this paper, we only consider $\ReLU$ function as the activation function: $\sigma(a)=\ReLU(a)=\max(0,a)$.

In this paper, we consider $L^{\infty}$ norm the max value of distance of each dimension, that is $d(\vx,\boldsymbol{x}_0)=\max |\boldsymbol{x}(n)-\boldsymbol{x}_0(n)|$. 
\fi

%\input{valabstraction}

\input{formula}

\input{Comparison}

\input{experiments}

\section{Conclusion}
{\color{blue}
	In this paper, we developed a novel Utility function to select few ReLU nodes to consider with binary variables to compute accurately bounds on neurons of DNNs. The novelties are that it focuses on {\em improvement} wrt a given input, rather than on generic {\em sensitivity} of a neuron wrt to a ReLU node, and it uses the solution of one call to an (efficient LP) solver to evaluate this improvement. This makes the choice particularly efficient, necessitating $\approx4$x less integer variables than previous proposals \cite{DivideAndSlide} for the same accuracy.} 
Our empirical studies revealed that this can yield highly accurate results, verifying up to 40\% more images than the SOTA ($\alpha,\beta$-Crown, winner of the 4 last VNNComp), with the same runtime, for DNNs with up to 20 000 neurons. 
The reason is that $\alpha,\beta$-Crown hits a complexity barrier, similarly as other competing solutions, when considering hard (even small) DNNs. This opens a lot of perspectives, among which: verifying efficiently other hard instances; certifying $\epsilon$-robustness of images for $\epsilon$ as large as possible; 
verifying global rather than local properties \cite{lipshitz}.

{\bf Reproducibility Statement:} 
We tested twice outlier results to confirm them, making sure of reproducibility on the given hardware. Precise details on the settings used are provided in the appendix.
Additional results {\color{blue}(e.g. ablation studies)} are also provided in the appendix.
Tested DNNs as well as MNIST and CIFAR10 DataSet are freely available.
The source code of Hybrid MILP will be provided on GitHub after acceptance (needing Gurobi as well as $\alpha,\beta$-Crown).

\newpage

\bibliography{references}
\bibliographystyle{plain}

\bigskip

\appendix

\input{proofsb}

\end{document}