### 2025-05-06 lipschitz/Lipitschz.tex
update the draft

\documentclass{llncs}
\usepackage{hyperref}
\usepackage{url}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{cases}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\vW}{\boldsymbol{W}}

\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}
\newcommand{\Improve}{\mathrm{Improve}}
\newcommand{\Utility}{\mathrm{SAS}}
\newcommand{\Sol}{\mathrm{Sol}}
\newcommand{\sol}{\mathrm{sol}}

\newcommand{\UB}{\mathrm{UB}}
\newcommand{\LB}{\mathrm{LB}}
\newcommand{\ub}{\mathrm{ub}}
\newcommand{\lb}{\mathrm{lb}}
\newcommand{\B}{\mathrm{B}}
\usepackage{amsmath, amssymb, amsfonts}

\newcommand{\ReLU}{\mathrm{ReLU}}

\newcommand{\CMP}{{\textrm{CMP}}\ }

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\toolname}{Hybrid MILP}

\title{Solution-aware vs global ReLU selection: \\
	partial MILP strikes back for DNN verification}
\date{}

\begin{document}

	\maketitle
	\linenumbers
	\begin{abstract}
		TODO
	\end{abstract}

	\section{Introduction}

	\section{Notations and Preliminaries}

	In this paper, we will use lower case latin $a$ for scalars, bold $\boldsymbol{z}$ for vectors, 
	capitalized bold $\boldsymbol{W}$ for matrices, similar to notations in \cite{crown}.
	To simplify the notations, we restrict the presentation to feed-forward, 
	fully connected ReLU Deep Neural Networks (DNN for short), where the activation function is $\ReLU : \mathbb{R} \rightarrow \mathbb{R}$ with
	$\ReLU(x)=x$ for $x \geq 0$ and $\ReLU(x)=0$ for $x \leq 0$, which we extend componentwise on vectors.

	%In this paper, we will not use tensors with a dimension higher than matrices: those will be flattened.

	%\subsection{Neural Network and Verification}

	% testtesttesttest
	An $\ell$-layer DNN is provided by $\ell$ weight matrices 
	$\boldsymbol{W}^i \in \mathbb{R}^{d_i\times d_{i-1}}$
	and $\ell$ bias vectors $\vb^i \in \mathbb{R}^{d_i}$, for $i=1, \ldots, \ell$.
	We call $d_i$ the number of neurons of hidden layer $i \in \{1, \ldots, \ell-1\}$,
	$d_0$ the input dimension, and $d_\ell$ the output dimension.

	Given an input vector $\boldsymbol{z}^0 \in \mathbb{R}^{d_0}$, 
	denoting $\hat{\boldsymbol{z}}^{0}={\boldsymbol{z}}^0$, we define inductively the value vectors $\boldsymbol{z}^i,\hat{\vz}^i$ at layer $1 \leq i \leq \ell$ with
	\begin{align*}
		\boldsymbol{z}^{i} = \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{i-1}+ \vb^i \qquad \, \qquad
		\hat{\boldsymbol{z}}^{i} = \ReLU({\boldsymbol{z}}^i).
	\end{align*}

	The vector $\hat{\boldsymbol{z}}$ is called post-activation values, 
	$\boldsymbol{z}$ is called pre-activation values, 
	and $\boldsymbol{z}^{i}_j$ is used to call the $j$-th neuron in the $i$-th layer. 
	For $\boldsymbol{x}=\vz^0$ the (vector of) input, we denote by $f(\boldsymbol{x})=\vz^\ell$ the output. Finally, pre- and post-activation neurons are called \emph{nodes}, and when we refer to a specific node/neuron, we use $a,b,c,d,n$ to denote them, and $W_{a,b} \in \mathbb{R}$ to denote the weight from neuron $a$ to $b$. Similarly, for input $\boldsymbol{x}$, we denote by $\val_{\boldsymbol{x}}(a)$ the value of neuron $a$ when the input is $\boldsymbol{x}$.

	For convenience, we write $n < z$ if neuron $n$ is on a layer before $\ell_z$, and $n \leq z$ if $n< z$ or $n=z$.

	%Concerning the verification problem, we focus on the well studied local-robustness question. Local robustness asks to determine whether the output of a neural network will be affected under small perturbations to the input. 
	%Formally, for an input $\vx$ perturbed by $\varepsilon >0$ under distance $d$, then the DNN is locally $\varepsilon$-robust in $\vx$ whenever:
	%\begin{align*}
	%	\forall \boldsymbol{x'} \text{ s.t. } d(\vx,\vx')\leq \varepsilon, \text{ we have }  
	%	argmax_i (f(\boldsymbol{x'})[i]) = argmax_i(f(\boldsymbol{x})[i])
	%\end{align*}

	\section{pMILP and Local robustness}

	\subsection{MILP and LP encodings for DNNs}

	Mixed Integer Linear Programming (MILP) value abstraction is a complete (but inefficient) method. 
	For an unstable neuron $n$ that its value $x \in [\LB(n),\UB(n)]$ with $\LB(n)<0<\UB(n)$, the value $\hat{x}$ of $\ReLU(x)$ can be encoded exactly in an MILP formula with one 
	integer variable $a$ valued in $\{0,1\}$, using constants $\UB(n),\LB(n)$ with 4 constraints \cite{MILP}:

	\vspace{-0.4cm}
	\begin{equation}\quad \hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0, \quad \wedge \quad \hat{x} \leq \UB(n) \cdot a \quad \wedge \quad \hat{x} \leq x-\LB(n) \cdot (1-a)
		\label{eq11}
	\end{equation}

	%For all $x \in [\LB(n),\UB(n)] \setminus 0$, there exists a unique solution $(a,\hat{x})$ that meets these constraints, with $\hat{x}=\ReLU(x)$ \cite{MILP}. The value of $a$ is 0 if $x < 0$, and 1 if $x>0$, and can be either if $x=0$. This encoding approach can be applied to every (unstable) ReLU node, and optimizing its value can help getting more accurate bounds. However, for networks with hundreds of {\em unstable} nodes, the resulting MILP formulation will contain numerous integer variables and generally bounds obtained will not be accurate, even using powerful commercial solvers such as Gurobi.

	MILP instances can be linearly relaxed into LP over-abstraction, where variables originally restricted to integers in $\{0,1\}$ (binary) are relaxed to real numbers in the interval $[0,1]$, while maintaining the same encoding. As solving LP instances is polynomial time, this optimization is significantly more efficient. However, this efficiency comes at the cost of precision, often resulting in less stringent bounds. This approach is termed the {\em LP abstraction}. We invoke a folklore result on the LP relaxation of (\ref{eq11}), for which we provide a direct and explicit proof:

	\subsection{partial MILP}

	We will use {\em partial MILP} (pMILP for short, see \cite{DivideAndSlide}) to get trade-offs between accuracy and runtime. Here, we extract the explanation from paper CITE.

	Let $X$ be a subset of the set of unstable neurons, and $n$ a neuron for which we want to compute upper and lower bounds on values: the partial MILP based on $X$ to compute neuron $n$ simply calls Gurobi to minimize/maximize the value of $n$ with the MILP model encoding, where variable $a$ is:
	\begin{itemize}
		\item binary for neurons in $X$ (exact encoding of the ReLU),
		\item linear for neurons not in $X$ (linear relaxation).
	\end{itemize}
	We will denote above case by MILP$_X$ or $\mathcal{M}_X$. Informally, we say that a node is opened if it is in $X$.

	If $X$ covers all unstable neurons, then MILP$_X$ is exact. But then the time cost may be very large (exponential growth).

	To reduce the runtime, we will limit the size of subset $X$. This a priori hurts accuracy. To recover some of this accuracy, we use an iterative approach: computing lower and upper bounds $\LB,\UB$ for neurons $n$ of a each layer iteratively, that are used when computing values of the next layer.

	\subsection{SAS}

	In pMILP, to decide the set $X$, we introduce the method {\em Solution-Aware Scoring} (SAS)
	to evaluate accurately how opening a ReLU impacts the accuracy. Again, here we use the definition from paper CITE. For details and explanation, see CITE.

	Assume that we want to compute an upper bound for neuron $z$ on layer $\ell_z$. For each node $n<z$, we denote ($\Sol\_\max_X^z(n))_{n \leq z}$ a solution of $\mathcal{M}_X$ maximizing $z$: $\Sol\_\max_X^z(z)$ is the maximum of $z$ under $\mathcal{M}_X$; and we denote $(\sol(n))_{n \leq z} = (\Sol\_\max_\emptyset^z(n))_{n \leq z}$ a solution maximizing the value for $z$ when all ReLU use the LP relaxation. Moreover,  we define the function
	$\Improve\_\max^z(n)=$ $\sol(z) - \Sol\_\max_{\{n\}}^z(z)$, 
	accurately represents how much opening neuron $n < z$ reduces the maximum computed for $z$
	compared with using only LP.

	First, SAS will call solvers to compute the LP model to get a solution, which is reasonably fast as there is no binary variables.

	Next, for a neuron $b$ on the layer before layer $\ell_z$, we define:

	\vspace{-0.4cm}
	$$\Utility\_\max\nolimits^z(b) = W_{bz} \times (\sol(\hat{b})- \ReLU(\sol(b)))$$
	\vspace{-0.4cm}

	And for a neuron $a$ two layers before $\ell_z$, 
	$b$ denoting neurons in the layer $\ell$ just before $\ell_z$.
	Recall the rate $r(b)=\frac{\max(0,\UB(b))}{\max(0,\UB(b))-\min(0,\LB(b))} \in [0,1]$.
	We define:

	\begin{flalign*}
		\Delta(\hat{a}) &= \ReLU(\sol(a))-\sol(\hat{a})&&\\
		\forall b \in \ell, \Delta(b) &= W_{ab}\Delta(\hat{a})&&\\	
	\end{flalign*}

	\vspace{-1.2cm}

	\begin{subnumcases}{\forall b \in \ell, \Delta(\hat{b}) =}
		r(b)\Delta(b), & for $W_{bz} > 0$ \\
		\max(\Delta(b),-\sol(b)), & for $W_{bz} < 0$ and $\sol(b)\geq0$\\
		\max(0,\Delta(b)+\sol(b)), & for $W_{bz} < 0$ and $\sol(b)<0$ \quad \, \quad \, \quad		 
	\end{subnumcases}

	\begin{flalign*}
		\Utility\_\max\nolimits^z(a) &= \Delta(z) = -\sum_{b \in \ell} W_{bz} \Delta(\hat{b})&&
	\end{flalign*}

	From paper CITE, we know that $\Utility$ is a safe overapproximation in the sense of following proposition:

	\begin{proposition}
		$0 \leq \Improve\_\max^z(a) \leq \Utility\_\max^z(a)$. 
	\end{proposition}

	\section{Global robustness and Lipschitz constant}

	\subsection{MILP and LP encodings for DNNs}

	At the other end of the spectrum, we find the Mixed Integer Linear Programming (MILP) value abstraction, which is a priori a complete method (albeit much slower than DeepPoly). 
	Consider an unstable neuron $n \in[\alpha,\beta]$. The value $x$ of $\ReLU(n)$ can be encoded exactly in an MILP formula with one integer (actually even binary) variable $a$ valued in ${0,1}$, using 4 constraints \cite{MILP}:
	\vspace{-0.1cm}
	\begin{align*}
		\hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0, \quad \wedge \quad \hat{x} \leq \beta \cdot a \quad \wedge \quad \hat{x} \leq x-\alpha \cdot (1-a)
	\end{align*}

	It can be shown that for all $x \in [\alpha,\beta] \setminus 0$, there exists a unique solution $(a,\hat{x})$ that meets these constraints, with $\hat{x}=\ReLU(x)$. Here, $a$ is 0 if $x < 0$, 1 if $x>0$, and can be either if $x=0$~\cite{MILP}. This encoding approach can be applied to every (unstable) ReLU node, and optimizing its value can help in certifying a given input. However, for networks with hundreds of nodes or more, the resulting MILP formulation will contain numerous integer variables and generally cannot be solved efficiently.

	MILP instances can be linearly relaxed into LP over-abstraction, where variables originally restricted to integers in ${0,1}$ (binary) are relaxed to real numbers in the interval $[0,1]$, while maintaining the same encoding. As solving LP instances is polynomial time, this optimization is significantly more efficient. However, this efficiency comes at the cost of precision, often resulting in less stringent bounds. This approach is termed the {\em LP abstraction}

	The following proposition shows that the LP abstraction refines the DeepPoly abstraction. 
	The difference between the LP and the DeepPoly abstractions is that LP simultaneous considers the two linear lower bounding functions for the ReLU operation, namely $ReLU(x) \geq 0$ and $ReLU(x) \geq x$, while the DeepPoly abstraction restricts itself to the application of a single one of these linear bounding functions.

	\begin{proposition}
		\label{LP}
		Given $x \in [\alpha,\beta]$ with $\alpha < 0 < \beta$, the following two systems of constraints 
		1) for LP and 2) for DeepPoly with both lower bounds are equivalent:
		\begin{align}
			& \hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0, \quad \wedge \quad \hat{x} \leq \beta \cdot a \quad \wedge \quad \hat{x} \leq x-\alpha \cdot (1-a), \, a \in [0,1]  \label{eq:lp}\\
			&\hat{x} \geq x \quad \wedge \quad \hat{x} \geq 0 \quad \wedge \quad \hat{x} \leq \beta \frac{x-\alpha}{\beta-\alpha} \label{eq:deeppoly}
		\end{align} 
	\end{proposition}

	\begin{proof}
		We need to show that the two upper bound constraints in Eq~\ref{eq:lp} is equivalent to the upper bound in Eq~\ref{eq:deeppoly}. To this end, we focus on the upper bound function in the linear variable $a \in  [0,1]$, with $x \in [\alpha,\beta]$ fixed. We have $\hat{x}$ in Eq~\ref{eq:lp} is upper bounded by $max_{a \in [0,1]} (min(\beta \cdot a, x - \alpha (1-a)))$, and this bound can be reached. Furthermore, 
		The function $\min(\beta \cdot a, x - \alpha (1-a))$ attains its maximum when $\beta \cdot a = x - \alpha (1-a)$, leading to the equation $(\beta - \alpha) a = x - \alpha$ and consequently $a = \frac{x - \alpha}{\beta-\alpha}$. This results in an upper bound of $\beta \cdot a = \beta \frac{x - \alpha}{\beta-\alpha}$ \qed
	\end{proof}

	\subsubsection*{Constraints for $\gamma$}

	We could get the following constraints for $\gamma$, using $y_i$ and $\hat{y}_i$ to denote the $x_i-x_i'$ and $\hat{x}_i-\hat{x}_i'$:\begin{align*}
		\hat{y}_i \leq a \gamma_i \hspace*{2ex} &\wedge \hspace*{2ex}\hat{y}_i \geq y_i - a \gamma_i\\
		\hat{y}_i \geq (a-1) \gamma_i  \hspace*{2ex} &\wedge \hspace*{2ex} \hat{y}_i \leq y_i + (1-a) \gamma_i,
	\end{align*} where $a$ is a binary variable, $\gamma_i$ is the upper bound of $x_i-x'_i$.

	The following plot will show the above constraints:

	\begin{tikzpicture}
		\begin{axis}[
			xlabel={$y_i$},
			ylabel={$\hat{y}_i$},
			xmin=-2, xmax=2,
			ymin=-2, ymax=2,
			axis lines=center,
			samples=100, 
			unit vector ratio=1 1 1, scale=1, xtick   = {-2,2},
			xticklabels = {$-\gamma_i$,$\gamma_i$},
			yticklabels = {},
			]
			\addplot[blue, thick, fill=blue, fill opacity=0.4] {x} \closedcycle; 
			\addplot[blue, thick] {0};

			\addplot[only marks, mark=*, mark size=2pt, blue] coordinates {(-2,-2)};
			\node[label={above:$(-\gamma_i,-\gamma_i)$}] at (axis cs: -1.35, -2.1) {};

			\addplot[only marks, mark=*, mark size=2pt, blue] coordinates {(2,2)};
			\node[label={above:$(\gamma_i,\gamma_i)$}] at (axis cs: 1.4, 1.5) {};
		\end{axis}
	\end{tikzpicture}

	The relation between $y_i$, $x'_i$ and $\hat{y}_i$ is $\hat{y}_i = \ReLU(x'_i+y_i)-\ReLU(x'_i).$

	We can use the following extra constraints (besides 4 constraints about $\hat{x}_i'=\ReLU(x_i')$) to build this relation:

	\begin{align*}
		&\hat{y}_i \geq -\hat{x}'_i \hspace*{1ex} \wedge \hspace*{1ex} \hat{y}_i \leq -\hat{x}'_i+a\beta_i  \hspace*{1ex}\wedge\hspace*{1ex} x_i'+y_i \leq a\beta_i \hspace*{1ex}\wedge\hspace*{1ex}  x_i'+y_i \geq (1-a)\alpha_i \\
		&\hat{y}_i \geq -\hat{x}'_i+(x_i'+y_i) \hspace*{1ex}\wedge\hspace*{1ex} \hat{y}_i \leq -\hat{x}'_i+(x_i'+y_i) +(a-1)\alpha_i \\
	\end{align*}

	Moreover, we can add two more natural constraints: $x_i'+y_i \geq \alpha_i \hspace*{1ex}\wedge\hspace*{1ex}  x_i'+y_i \leq \beta_i.$

	A more direct way to establish the constraints is as follows: \begin{align*}
		& y_i+x'_i \leq a\beta_i \quad\wedge \quad y_i+x'_i\geq (1-a)\alpha_i\\	
		& x_i' \leq a'\beta_i \quad\wedge \quad x_i'\geq (1-a')\alpha_i\\
		&\hat{y}_i \leq a\gamma_i \quad\wedge \quad	\hat{y}_i \geq -a'\gamma_i \\
		&	\hat{y}_i \leq y_i+(1-a)\gamma_i \quad\wedge \quad	\hat{y}_i \geq y_i - (1-a')\gamma_i \\
		&	\hat{y}_i \leq -x'_i+a\beta_i \quad\wedge \quad	\hat{y}_i \geq -x'_i+(1-a')\alpha_i \\
		&	\hat{y}_i \leq y_i+x'_i+(1-a)(-\alpha_i)\quad\wedge \quad	\hat{y}_i \geq y_i+x'_i+a'(-\beta_i) \\
	\end{align*}The following two plots show its shape:

	\iffalse
	\hspace*{-10ex}
	\begin{tikzpicture}
		\begin{axis}[	axis on top, xlabel = \(x'_i\),
			ylabel = {\(y_i\)}, zlabel = \(\hat{y}_i\),
			set layers=default,
			xmax = 4, xmin = -4,
			ymax = 1, ymin = -1,		
			zmax = 1, zmin = -1,
			unit vector ratio=1 1 1, scale=3,
			view={60}{50}, ytick   = {-1,0,1},
			yticklabels = {$-\gamma_i$,$0$,$\gamma_i$}, xtick = {0},
			xticklabels = {$0$}, ztick   = {-1,0,1},
			zticklabels = {$-\gamma_i$,$0$,$\gamma_i$},
			]
			\addplot3[fill=blue,opacity=0.1, fill opacity=0.4] 
			coordinates {
				(0,0,0) (-1,1,0) (-4,1,0) (-4,-1,0) (0,-1,0) (0,0,0)
			};

			\addplot3[fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0) (0,1,1) (4, 1, 1) (4, -1, -1) (1,-1,-1) (0,0,0)
			};

			\addplot3[fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0)  (-1,1,0) (0,1,1) (0,0,0)
			};

			\addplot3[fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0)  (0,-1,0) (1,-1,-1) (0,0,0)
			};

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(1,-1,-1)};
			\node[label={$(\gamma_i,-\gamma_i, -\gamma_i)$}] at (axis cs: 1.2, -0.5 ,-1) {};

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(-1,1,0)};
			\node[label={$(-\gamma_i,\gamma_i, 0)$}] at (axis cs: -1, 0.8 ,0) {};

		\end{axis}
	\end{tikzpicture}
	\fi

	\hspace*{-15ex}
	\begin{tikzpicture}
		\begin{axis}[	axis on top, xlabel = \(x'_i\),
			ylabel = {\(y_i\)}, zlabel = \(\hat{y}_i\),
			set layers=default,
			xmax = 4, xmin = -4,
			ymax = 1, ymin = -1,		
			zmax = 1, zmin = -1,
			unit vector ratio=1 1 1, scale=2.5,  ytick   = {-1,0,1},
			yticklabels = {$-\gamma_i$,$0$,$\gamma_i$}, xtick = {0},
			xticklabels = {$0$}, ztick   = {-1,0,1},
			zticklabels = {$-\gamma_i$,$0$,$\gamma_i$},
			view={35}{14},
			]
			\addplot3[ fill=blue,opacity=0.1, fill opacity=0.4] 
			coordinates {
				(0,0,0) (-1,1,0) (-4,1,0) (-4,-1,0) (0,-1,0) (0,0,0)
			};

			\addplot3[	fill=blue,opacity=0.1, fill opacity=0.4] 
			coordinates { (0,0,0) (0,1,1) (4, 1, 1) (4, -1, -1) (1,-1,-1) (0,0,0)
			};

			\addplot3[	fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0)  (-1,1,0) (0,1,1) (0,0,0)
			};

			\addplot3[	fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0)  (0,-1,0) (1,-1,-1) (0,0,0)
			};

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(1,-1,-1)};
			\node[label={$(\gamma_i,-\gamma_i, -\gamma_i)$}] at (axis cs: 1.2, -0.5 ,-1) {};

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(-1,1,0)};
			\node[label={$(-\gamma_i,\gamma_i, 0)$}] at (axis cs: -1, 0.8 ,0) {};

		\end{axis}
	\end{tikzpicture}

	\section{Conclusion}

	TODO

	\bibliography{references}
	\bibliographystyle{plain}

	\bigskip

	\appendix

\end{document}