### 2025-07-16 ATVA 2025/Final_version/Lipschitz.tex
update the final version

\section{Lipschitz}

We will introduce some formulas.

\subsection*{GS formula}

$$v=\frac{\gamma}{4}\cdot\lambda_a\cdot\sign(\lambda_a).$$

Here, $\lambda_a$ is computed by backward-propagation as usual (the formula at the end of page 9), but use $-\gamma$ as $\LB$ and $\gamma$ as $\UB$.

\subsection*{SAS formula}

 $$v = \Delta(\hat{a})\cdot\lambda_a.$$

Here, the computation of $\Delta(\hat{a})$ is as follows:

\begin{subnumcases}{\Delta(\hat{a}) =}
\sol(a)-\sol(\hat{a}), &for $\lambda_a > 0$ and $\sol(a) \geq 0$\\
-\sol(\hat{a}), &for $\lambda_a > 0$ and $\sol(a) \leq 0$\\
-\sol(\hat{a}), &for $\lambda_a < 0$ and $\sol(a) \geq 0$\\
\sol(a)-\sol(\hat{a}), &for $\lambda_a < 0$ and $\sol(a) \leq 0$	 
\end{subnumcases}