### 2025-07-16 ATVA 2025/Final_version/formula_part2.tex

\section{Three or More Layer before  $z$.}

Suppose $a$ is a neuron in three or more layers before $z$, we use $b$ to denote neurons in the next layer after $a$ and $c$ to denote neurons in two layers after $a$ and so on.

This formula is based on previous subsection but more complex. In some network, running this formula may cost too much time.

\begin{definition}\label{3layer}
Let $\UB$ and $\LB$ denote the precomputed upper bounds and lower bounds used in building MILP models. We define the following function $h$ for all neurons $b$ in layers between $a$ and $z$ as follows:
	\begin{align}
		&v_0 = sol(\hat{b}), v_1 = \ReLU(sol(b)), v_2 = \frac{\UB(b)sol(b)-\UB(b)\LB(b)}{\UB(b)-\LB(b)}\\
		&h(b) =
		\begin{cases}
			\frac{v_0-v_1}{v_2-v_1}, & \text{if } v_2-v_1 > 0\\
			0.5, & \text{otherwise.}
		\end{cases}
	\end{align} 
\end{definition}

\begin{definition}
	Continue the assumption in Definition \ref{3layer}. We define function $D$ layer by layer.

	First, $D(a) = \ReLU(sol(a))-sol(\hat{a})$.

Next we define all neurons $b$ in one layer after $a$: \begin{align}
	&u_0 = \max(\LB(b),\min(\UB(b),  sol(b)+D(a)W_{ab}))\\
	&u_1 = \begin{cases}
		\ReLU(u_0)+h(b)(\frac{\UB(c)u_0-\UB(b)\LB(b)}{\UB(b)-\LB(b)}-\ReLU(u_0)), & \text{if }\LB(b) < 0\\
	u_0, & \text{if }  \LB(b) \geq 0
	\end{cases}\\
	&D(b) = u_1-sol(\hat{b})
\end{align}

	Suppose $y$ is a neuron in at least one layer before $z$, and we have defined $D(x)$ for all neurons $x$ in layers before $y$, we define $D(y)$ as follows:
	\begin{align}
		&w_0 = \sum_x D(x)W_{xy}\\
		&w_1 = \min(\UB(y),sol(y)+w_0)\\		
		&D(y) =
		\begin{cases}
			w_1-sol({y}), & \text{if }W_{yz} > 0 \text{ and } \LB(y)\geq 0\\
		k(y)(w_1-sol({y})), & \text{if }W_{yz} > 0 \text{ and } \LB(y)< 0\\
		\ReLU(w_1)-sol(\hat{y})	, & \text{if }  W_{yz} < 0
		\end{cases}\\
	\end{align}	Finally, 	\begin{align*}
	&D(z) = \sum_y D(y)W_{yz}\\
	&Utility\_max^z(a) = -D(z)
	\end{align*}

\end{definition}