### 2025-07-28 AAAI2026/introduction.tex

Deep neural networks (DNNs for short) have demonstrated remarkable capabilities, achieving human-like or even superior performance across a wide range of tasks. However, their robustness is often compromised by their susceptibility to input perturbations \cite{szegedy}. This vulnerability has catalyzed the verification community to develop various methodologies, each presenting a unique balance between completeness and computational efficiency \cite{Marabou,Reluplex,deeppoly}. This surge in innovation has also led to the inception of competitions such as VNNComp \cite{VNNcomp}, which aim to systematically evaluate the performance of neural network verification tools. Among them, NNenum \cite{nnenum}, Marabou \cite{Marabou,Marabou2}, and PyRAT \cite{pyrat} MnBAB \cite{ferrari2022complete}, built upon ERAN \cite{deeppoly} and PRIMA \cite{prima}; and importantly, $\alpha,\beta$-CROWN \cite{crown,xu2020fast}, the winner of the last 4 VNNcomp, benefiting from branch-and-bound based methodology \cite{cutting,BaB}.

The benchmarks usually focus on {\em local} robustness, i.e. given a DNN, an image and a small neighborhood around this image, is it the case that all the images in the neighborhood are classified in the same way by the DNN? The neighborhood is provided by a maximal perturbation of the input image, often an $L_0$ perturbation, i.e. every subpixel of the input image can vary in a very small range, typically 
$\frac{2}{255}$ (that is 2 levels of grey/blue/right/green), as $L_0$ is perfectly linear and perturbations of subpixels are independent. While the $L_0$ norm is the easiest to handle, it is not necessarily the most meaningful perturbation.
More importantly, verification tools for local robustness are hard to be used in a real-time decision making pipeline: considering an autonomous car with a video feed from the dashboard, it is computationally too intensive to certify in few ms images of the feed to e.g. skip non robust images (or operate in a degraded mode - e.g. driving safely) and only consider certified robust ones for the decision-making process.

\smallskip

In this paper, we consider {\em global} robustness, that is we do not restrict the certification process to the neighborhood of an input. We propose a two steps procedure. The first step, performed offline once, 
computes global bounds on the shift between 
output values of different decision classes due to the perturbation. That is, considering decision classes $A$ and $B$ and perturbation $\varepsilon$, compute 
{\em global bound} $B^\varepsilon_{A,B} \geq \max_{I,I', |I-I'| \leq \epsilon}(value_{I}(A) - value_{I}(B) + value_{I'}(B) - value_{I'}(A))$, where $value_{J}(X)$ is the output value of class $X \in \{A,B\}$ for input image $J \in \{I,I'\}$. Notice that as 
$I,I',A,B$ have symetrical roles, we can choose $B^\varepsilon_{A,B} = B^\varepsilon_{B,A}$. Also, as $I,I'$ have symetrical role, the minimum value is exactly $-$ the maximum value. 
%Hence, if we have $n$ decision classes, we only have to compute $n (n-1)/2$ bounds. 
Bound $B^\varepsilon_{A,B}$ needs to be computed once for a DNN, 
and it is valid over the whole input space; compared with $k$ computations for {\em local robustness}, once for each of the $k$ input images, with results only valid for these $k$ images.

The second step is real-time, being performed with the DNN inference of the image $I$ considered: it suffices to consider the class $A$ with the highest output value $value_{I}(A)$, and check whether for every other class $B \neq A$, 
$value_{I}(A) - value_{I}(B) > B^\varepsilon_{A,B}$. 
If this is the case, then we are certified that image $I$ is robust for perturbation $\varepsilon$, because $\varepsilon$-perturbed image $I'$ could at most get  $value_{I'}(B) \leq B^\varepsilon_{A,B}  - (value_{I}(A) - value_{I}(B))  + value_{I'}(A) < value_{I'}(A)$, hence $A$ is also the predicted class for image $I'$. Otherwise, one could either skip image $I$ (in a video feed), or use safer degraded mode in the decision making process till a more trustable robust image is received.

Our main contributions address the challenges to compute the {\em global bounds} $B^\varepsilon(A,B)$, for $A,B$ output neurons of a DNN:
\begin{enumerate}
	%\item  Our first contribution studies the {\em LP relaxation} of the exact MILP encoding of ReLUs. {\color{blue} We establish in Proposition \ref{LP} its equivalence with the so-called "triangular abstraction"}.

	\item We revisit the idea from \cite{DivideAndSlide} to consider small calls to a partial MILP (pMILP) solver, i.e. with few binary variables encoding few ReLU functions exactly (other being encoded with more efficient but less accurate linear variables), to compute bounds for each neuron inductively, hence many ($O(n)$, the number of neuron) small calls, with a complexity exponential only in the few binary variables (Section 4). Compared to the few (one per output neuron) complex call to BaB, each with a worst case complexity exponential in the number of neurons of the DNN (which is far from the actual complexity thanks to pruning branches in BaB - but which can be too large as we shown in the 6 first DNNs of Table \ref{table_beta}). Two questions arise: how to select few very important ReLUs?, and is computing the bounds of intermediate neurons a good trade-off compared with the theoretical loss of accuracy due to selecting only some binary ReLUs? Answer to these questions were not looking very promising judging by previous attempt \cite{DivideAndSlide}, which was using a simple selection heuristic of nodes in the previous layer only.

	\item On the first question, we adapted from BaB-SR \cite{BaB} and FSB \cite{FSB}, which choose branching nodes for BaB, {\em global scoring ({\sf GS})} functions to choose important ReLUs (Section 5). These revealed to be much more accurate than the simple heuristic in 
	\cite{DivideAndSlide}. However, we also uncover that the {\em improvement function} that {\sf GS} tries to approximate depends heavily upon the mode of the ReLU function, and as this mode is unavailable to {\sf GS}, there are many cases in which {\sf GS} is far from the improvement (with both under or over-approximation of the improvement function).

	\item We thus designed a {\em novel solution-aware scoring ({\sf SAS})}, which uses the solution of a unique LP call, that provides the mode to consider (Section 6). Theoretically, we show that {\sf SAS} is always an over-approximation of the improvement (Proposition \ref{prop2}), which implies that a small {\sf SAS} value implies that the ReLU is unnecessary. Experimentally, we further show that {\sf SAS} is very close to the actual improvement, closer than {\sf GS}, and that overall, the accuracy from SAS is significantly better than {\sf GS}.  Compared with the heuristic in \cite{DivideAndSlide},  {\sf SAS} is much more efficient ($\approx 6$ times less binary variables for same accuracy (Fig. \ref{fig_table3})). 
	%We however explain why focusing on a solution may not be appropriate to choose branching order within this selection, and that GS would be better suited, which we verify experimentally (see Table. \ref{table.order}).

	\item Compared with many calls using full MILP, where all the ReLUs are encoded as binary variables, {\sf SAS} (and {\sf GS}) encode only a subset of ReLUs as binary and others as linear variables. The model in full MILP is fully accurate, while {\sf SAS} (and {\sf GS})
	are abstractions, and thus much faster to solve. 
	While full MILP is thus {\em asymptotically} more accurate than {\sf SAS} and {\sf GS}, {\em experimentally}, every reasonable time-out leads to much better practical accuracy of {\sf SAS} (and {\sf GS}) (see Fig. \ref{fig555}).

	%	\item {\color{blue} We designed a novel Utility function to choose few neurons to encode with the exact MILP encoding, while others are treated with the efficient LP relaxation, giving rise to partial MILP (pMILP). Specifically, the novelty of Utility resides in the use of the solution to an (efficient LP) solver on the node $z$ we want to bound. Utility can then precisely evaluate how much accuracy is gained by switching neuron $a$ from LP (solution of the LP call) to the exact MILP encoding of ReLU (exact computation from the solution, which can be made thanks to Proposition \ref{LP}), with a proved bound on the precision (Proposition \ref{prop2}). Because pMILP focuses on the {\em improvement} (binary - linear), it is much more efficient ($\approx 4$ times less integer variables for same accuracy (Table \ref{tab:example1})) than previous attempts, which consider the generic {\em sensitivity} to this neuron. To the best of our knowledge, this is the first time such a solution of an (LP) call is used to evaluate the contribution of each neuron, including heuristics for BaB, e.g. \cite{BaB,FSB}.}

	\item For the second question, we propose a new verifier, called {\em Hybrid MILP}, invoking first 	$\alpha,\beta$-CROWN with short time-out to settle the easy instances. On those ({\em hard}) instances which are neither certified nor falsified, we call pMILP with few neurons encoded as binary variables. Experimental evaluation reveals that Hybrid MILP achieves a beneficial balance between accuracy and completeness compared to prevailing methods. It reduces the proportion of undecided inputs from $20-58\%$ ($\alpha,\beta$-CROWN with 2000s TO) to $8-15\%$, while taking a reasonable average time per instance ($46-420$s), Table \ref{table_hybrid}. It scales to fairly large networks such as CIFAR-10 CNN-B-Adv \cite{SDPFI}, with more than 2 million parameters.
%We verify experimentally that the algorithm offers interesting trade-offs, by testing on local robustness for DNNs trained "naturally" (and thus difficult to verify).
%KSM: I think this is a distraction in intro, so I suggest moving to later part
% Overall, the worst case complexity of algorithm \ref{algo1} is lower than $O(N 2^K LP(N))$, where $N$ is the number of nodes of the DNN, $K$ the number of ReLU nodes selected as binary variable, and $LP(N)$ is the (polynomial time) complexity of solving a linear program representing a DNN with $N$ nodes. This complexity is an upper bound, as e.g. Gurobi is fairly efficient and never need to consider all of the $2^K$ ReLU configurations to compute the bounds. Keeping $K$ reasonably low thus provides an efficient algorithm. 
%By design, it will never run into a complexity wall (unlike the full MILP encoding), although it can take a while on large networks because of the linear factor $N$ in the number of nodes.
\end{enumerate}

Limitation: We consider DNNs employing the standard ReLU activation function, though our findings can be extended to other activation functions, following similar extention by \cite{DivideAndSlide}, with updated MILP models e.g. for maxpool.

%\newpage

%   
% 
%
%In this context, application of DNNs in safety critical applications is cautiously envisioned. For that to happen at a large scale, hard guarantees should be provided \cite{certification}, through e.g. incremental verification \cite{incremental}, so that to avoid dramatic consequences. It is the reason for the development of (hard) verification tools since 2016, with now many tools with different trade-offs from exact computation but slow (e.g. Marabou \cite{katz2019marabou}/Reluplex\cite{Reluplex}), up to very efficient but also incomplete (e.g. ERAN-DeepPoly \cite{deeppoly}). To benchmark these tools, a competition has been run since 2019, namely VNNcomp \cite{VNNcomp}. The current overall better performing verifier is $\alpha$-$\beta$-CROWN \cite{crown}, a fairly sophisticatedly engineered tool based mainly on "branch and bound" (BaB) \cite{BaB}, and which can scale all the way from complete on smaller DNNs \cite{xu2020fast} up to very efficient on larger DNNs, constantly upgraded, e.g. \cite{cutting}. 
%
%While the verification engines are generic, the benchmarks usually focus on local robustness, i.e. given a DNN, an image and a small neighbourhood around this image, 
%is it the case that all the images in the neighbourhood are classified in the same way.
%While some quite large DNNs (e.g. ResNet with tens of thousands of neurons) can be verified very efficiently (tens of seconds per input) \cite{crown}, with all inputs either certified robust or an attack on robustness is found; some smaller DNNs (with hundreds of neurons, only using the simpler ReLU activation function) cannot be analysed fully, with $12-20\%$ of inputs where neither of the decisions can be reached (\cite{crown} and Table \ref{tab:example}). Actually, DNNs which are trained to be robust (using DiffAI \cite{DiffAI} or PGD \cite{PGD}) are easier to verify, while the DNNs trained in a "natural" way are harder to verify.
%
%
%In this paper, we focus on DNNs trained in a "natural" way,
%%uncovering what makes the DNNs trained in a natural way so hard to verify (
%because for "easier" DNNs, adequate methods already exist. 
%To do so, we analyse the abstraction mechanisms at the heart of several efficient algorithms, namely Eran-DeepPoly \cite{deeppoly}, the Linear Programming approximation \cite{MILP}, PRIMA \cite{prima}, and different versions of ($\alpha$)($\beta$)-CROWN \cite{crown}. All these algorithms compute lower or/and upper bounds for the values of neurons (abstraction on values) for inputs in the considered input region, and conclude based on such bounds. For instance, if for all image $I'$ in the neighbourhood of image $I$, we have $weight_{I'}(n'-n) < 0$ for $n$ the output neuron corresponding to the expected class, then we know that the DNN is robust in the neighbourhood of image $I$. We restrict the formal study to DNNs using only the standard ReLU activation function, although nothing specific prevents the results to be extended to more general architectures. We uncover that {\em compensations} 
%(see next paragraph) is the phenomenon creating inaccuracies. We verified experimentally that a DNN trained in a natural way has heavier compensating pairs than DNNs trained in a robust way.
%
%Formally, a compensating pair is a pair of paths $(\pi,\pi')$ between a pair of neurons $(a,b)$, such that we have $w < 0 < w'$, for $w,w'$ the products of weight seen along $\pi$ and $\pi'$. Ignoring the (ReLU) activation functions, the weight of $b$ is loaded with $w \cdot weight(a)$ by $\pi$, while it is loaded with $w' \cdot weight(a)$ by $\pi'$. That is, it is loaded by $(w+w') weight(a)$. As $w,w'$ have opposite sign, they will compensate (partly) each other. The compensation is only partial due to the ReLU activation seen along the way of $\pi$ which can "clip" a part of $w \cdot weight(a)$, and similarly for $\pi'$. However, it is very hard to evaluate by how much without explicitly considering both phases of the ReLUs, which all the efficient tools try to avoid because it is very expansive (could be exponential in the number of such ReLU nodes opened).

%Our first main contribution is to formally show, in Theorem \ref{th1}, that compensation is the sole reason for the inaccuracies as (most) efficient algorithms will compute exact bounds for all neurons if there is no compensating pair of paths at all.
%While this theorem is theoretically interesting, it is not usable in practice as (almost) all networks have some compensating pairs. However, this notion of compensating pairs opens a first interesting idea concerning an exact abstraction of the network using a Mixed Integer Linear Program \cite{MILP}, where the weight of each neuron is a linear variable, and ReLU node may be associated with binary variables (exact encoding) or linear variables (overapproximation). While LP tools can scale to thousands of linear variables, MILP encoding can only be solved for a limited number of binary variables. This suggests that a simpler encoding could be used for those ReLUs that are not on compensating pairs, as their precise outcome may not be necessary.

%Our second main contribution is to show formally in Theorem \ref{th2}, that 
%encoding all ReLU nodes on a pair of compensating paths with a binary variable,
%and using linear relaxation for the other ReLU nodes, will lead to exact bounds for (most) of the algorithms considered. This theorem allows to restrict the number of integer variables, and thus to obtain encodings that are faster to solve. Practically, however, (almost) all ReLU nodes are on some compensating path, and using this exact restricted MILP encoding will be too time consuming.

%Our third main contribution is more practical, proposing Algorithm \ref{algo1} based on this knowledge that compensating pair of paths are the reason for inaccuracy. The idea is thus to use this information to rank the ReLU nodes in terms of importance, and only keep the most important ones as binary variables, and use linear relaxation for the least important ones.
%%More precisely, the algorithm will, as DeepPoly, consider layers one by one and neurons $b$ %on this layer one by one, selecting the heaviest pairs of compensating paths ending in $b$
%%and associating these nodes with a binary variable. Then an MILP tool such as Gurobi is used %to compute the lower and upper bound for node $b$. 
%Overall, the worst case complexity of algorithm \ref{algo1} is lower than $O(N 2^K LP(N))$, where $N$ is the number of nodes of the DNN, $K$ the number of ReLU nodes selected as binary variable, and $LP(N)$ is the (polynomial time) complexity of solving a linear program representing a DNN with $N$ nodes. This complexity is an upper bound, as e.g. Gurobi is fairly efficient and never need to consider all of the $2^K$ ReLU configurations to compute the bounds. Keeping $K$ reasonably low thus provides an efficient algorithm. 
%By design, it will never run into a complexity wall (unlike the full MILP encoding), although it can take a while on large networks because of the linear factor $N$ in the number of nodes. An additional interesting point is that it is extremely easy to parallelize, as all the nodes in the same layer can be run in parallel. We verify experimentally that the algorithm offers interesting trade-offs, by testing on local robustness for DNNs trained "naturally" (and thus difficult to verify).

%KSM: I suggest we move this to experimental evaluation
%This paper does not focus on producing the most efficient tool, and we did not spend engineering efforts to optimize it. The focus is instead on the novel notion of compensation, the associated methodology and its evaluation. For instance, our implementation is fully in Python, with uncompetitive runtime for our DeepPoly implementation ($\approx 100$ slower than in CROWN). Still, evaluation of the methodology versus even the most efficient tools reveals a lot of potential for the notion of compensation, opening up several opportunities for applying it in different contexts of DNN verification (see Section \ref{Discussion}).