### 2025-07-29 AAAI2026/global.tex
global

	\section{Modeling for global verification}

	\subsection{Extending the classical encoding}

    The most straightforward idea to encode the global verification question is to 
    consider variable $x_j,\hat{x}_j$ and $x'_j,\hat{x}'_j$ for nodes of the image $I$ and of its perturbation $I'$,
    using the classical encoding (\ref{eq11}) \cite{MILP} for each variable, 
    that we call $\mathcal{M}^{classical}$.
    Further:
    \begin{enumerate}
		\item For each input nodes $x_j,x'_j$ of $I,I'$, we add constraints
        $|x_j-x'_j| \leq \varepsilon$ 
        ($L_\infty$-perturbation - see next section for $L_1$).
		\item We set the optimization objective as maximize $x_i - x_j +x'_j - x'_i$ to compute $\beta^\varepsilon_{i,j}=\beta^\varepsilon_{j,i}$.
	\end{enumerate}

	This model contains twice as many (binary) variables as for local robustness. As the worst-case complexity of solving an MILP model 
    is exponential in the number of binary variables, it will be computationally intractable to compute $\beta^\varepsilon_{i,j}$ in this way, but for very small DNNs. As explained in the previous section, the standard way is to rely on relaxing at least some of the binary variables into linear variables.

    The problem is that the LP relaxation of this classical encoding is extremely inaccurate, as variables $x_j,x'_j$ dependencies are due solely on the function definition through common ancestor variables in previous layers - as this is approximated with LP, most of the dependencies will be lost.

    On top of being an issue with partial MILP models, LP relaxation is also used during the Branch and Bound process internal to MILP solvers to provide bounds for each branch. So even when all variables are opened and the MILP model is fully accurate, the classical encoding will likely not be efficient as well.

    We will check that in table \ref{table.classical} in Section Experimental Evaluation.

	\subsection{A novel "2v" model with {\em diff variables}}

    The idea to circumvent the loss of dependencies when applying linear relaxation is to switch the variables towards $x'_j,\hat{x}'_j, y_j, \hat{y}_j$, where $y_j, \hat{y}_j$ are the {\em diff variables}, with
    $y_j = x_j - x'_j$ and $\hat{y}_j= \hat{x}_j - \hat{x}'_j$. Intuitively, the diff variables will be small as they will be carrying the difference between both the image $I$ and its perturbation $I'$.
    We will use the classical MILP encoding for $\hat{x}'_j$.
    However, we cannot apply the classical MILP encoding (\ref{eq11}) \cite{MILP} on the diff variable $\hat{y}_j$. Instead we study the function of 
    $\hat{y}_j$ given $y_j$ and $x'_j$, and encode it into a novel MILP model.
    A first important observation is that $I,I'$ being perfectly symetrical,
    for $\gamma_i$ being a upper bound of $y_i$, then $-\gamma_i$ is a lower bound 
    for $y_i$.

	%	(This model has the same binary set, although the meaning of binary variable for $y_i$ is somehow different.)

	%	The exact constraints for $$ \begin{align*}
		%		&\hat{y}_i \geq -\hat{x}'_i \hspace*{1ex} \wedge \hspace*{1ex} \hat{y}_i \leq -\hat{x}'_i+a\beta_i  \hspace*{1ex}\wedge\hspace*{1ex} x_i'+y_i \leq a\beta_i \hspace*{1ex}\wedge\hspace*{1ex}  x_i'+y_i \geq (1-a)\alpha_i \\
		%		&\hat{y}_i \geq -\hat{x}'_i+(x_i'+y_i) \hspace*{1ex}\wedge\hspace*{1ex} \hat{y}_i \leq -\hat{x}'_i+(x_i'+y_i) +(a-1)\alpha_i \\
		%	\end{align*} 
	%	
	%	
	%	Moreover, we can add two more natural constraints: $x_i'+y_i \geq \alpha_i \hspace*{1ex}\wedge\hspace*{1ex}  x_i'+y_i \leq \beta_i.$

	We first display on Figure \ref{fig.2v} the function $f(x'_i,y_i)=\hat{y}_i$, when we know that $y_i \in [- \gamma_i, \gamma_i]$. When $x'_i < -\gamma_i$, then $x_i <0$ and we have $\hat{y}_i = \ReLU(x_i) - \ReLU(x'_i) = 0-0=0$.
    For $x'_i > \gamma_i$, we have $x_i >0$ and $\hat{y}_i=y_i$. Between these two linear planes, we have two more linear planes presenting intermediate cases, so 4 linear planes in total.

	\begin{figure}[b!]
		\centering
	\hspace*{-5ex}
	\begin{tikzpicture}[scale=0.65]
		\begin{axis}[	axis on top, xlabel = \(x'_i\),
			ylabel = {\(y_i\)}, zlabel = \(\hat{y}_i\),
			set layers=default,
			xmax = 4, xmin = -4,
			ymax = 1, ymin = -1,		
			zmax = 1, zmin = -1,
			unit vector ratio=1 1 1, scale=2.5,  ytick   = {-1,0,1},
			yticklabels = {$-\gamma_i$,$0$,$\gamma_i$}, xtick = {0},
			xticklabels = {$0$}, ztick   = {-1,0,1},
			zticklabels = {$-\gamma_i$,$0$,$\gamma_i$},
			view={35}{14},
			]
			\addplot3[ fill=blue,opacity=0.1, fill opacity=0.4] 
			coordinates {
				(0,0,0) (-1,1,0) (-4,1,0) (-4,-1,0) (0,-1,0) (0,0,0)
			};

			\addplot3[	fill=blue,opacity=0.1, fill opacity=0.4] 
			coordinates { (0,0,0) (0,1,1) (4, 1, 1) (4, -1, -1) (1,-1,-1) (0,0,0)
			};

			\addplot3[	fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0)  (-1,1,0) (0,1,1) (0,0,0)
			};

			\addplot3[	fill=blue,opacity=0.1, fill opacity=0.4	] 
			coordinates { (0,0,0)  (0,-1,0) (1,-1,-1) (0,0,0)
			};

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(1,-1,-1)};
			\node[label={$(\gamma_i,-\gamma_i, -\gamma_i)$}] at (axis cs: 1.2, -0.5 ,-1) {};

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(-1,1,0)};
			\node[label={$(-\gamma_i,\gamma_i, 0)$}] at (axis cs: -1, 0.8 ,0) {};

		\end{axis}
	\end{tikzpicture}
	\caption{The outcome $\val(\hat{y}_i)=\ReLU(\val(x_i))-\ReLU(\val(x'_i))$ 
	depending on $\val(y_i) = \val(x_i)-\val(x'_i)$ and on $\val(x'_i)$.}
	\label{fig.2v}
\end{figure}

Assuming that $y_i \in [-\gamma_i, \gamma_i]$,
and that $x'_i \in [\alpha_i,\beta_i]$,
we have that $\hat{y}_i = \ReLU(x'_i+y_i)-\ReLU(x'_i)$ is the solution of:
	\begin{align*}
		& \begin{aligned}
			y_i + x'_i &\leq a\beta_i        &
			y_i &+ x'_i \geq (1-a)\alpha_i \\
			x'_i       &\leq a'\beta_i       & 
			x'_i       &\geq (1-a')\alpha_i \\
			\hat{y}_i  &\leq a\gamma_i       &
			\hat{y}_i  &\geq -a'\gamma_i \\
			\hat{y}_i  &\leq y_i + (1-a)\gamma_i  &
			\hat{y}_i  &\geq y_i - (1-a')\gamma_i \\
			\hat{y}_i  &\leq -x'_i + a\beta_i &
			\hat{y}_i  &\geq -x'_i + (1-a')\alpha_i \\
			\hat{y}_i  &\leq y_i + x'_i + (1-a)(-\alpha_i) &
			\hat{y}_i  &\geq y_i + x'_i + a'(-\beta_i)
		\end{aligned}
	\end{align*} Here, $a,a'$ are binary variables, and $a'$ is also the binary variable in the constraints for $\hat{x}'_i=\ReLU(x'_i)$.

Similar to the second model, the constraints on $x_i$ for $\hat{x}_i=\ReLU(x_i)$ are not necessary, or at least can be relaxed without any loss in accuracy.

	%	\begin{align*}
		%		& y_i+x'_i \leq a\beta_i \quad\wedge \quad y_i+x'_i\geq (1-a)\alpha_i\\	
		%		& x_i' \leq a'\beta_i \quad\wedge \quad x_i'\geq (1-a')\alpha_i\\
		%		&\hat{y}_i \leq a\gamma_i \quad\wedge \quad	\hat{y}_i \geq -a'\gamma_i \\
		%		&	\hat{y}_i \leq y_i+(1-a)\gamma_i \quad\wedge \quad	\hat{y}_i \geq y_i - (1-a')\gamma_i \\
		%		&	\hat{y}_i \leq -x'_i+a\beta_i \quad\wedge \quad	\hat{y}_i \geq -x'_i+(1-a')\alpha_i \\
		%		&	\hat{y}_i \leq y_i+x'_i+(1-a)(-\alpha_i)\quad\wedge \quad	\hat{y}_i \geq y_i+x'_i+a'(-\beta_i) \\
		%	\end{align*}

	The following plot illustrates the constraints described above.

	In practice, when we relax the constraints for a node, we will convert $a_i$ and $a_i'$ to continuous variables together. In principle, we can also choose to change  only one of those two binary variables.

	%	The relaxation of this model is similar: let $a$s and $a'$s be continuous variables instead of binary/integer variables. Unlike the first model in this section, relaxing a few nodes does not lose too much accuracy.

	A special relaxation approach of this model is to treat the binary variables $a'$ appearing in the above constraints and the binary variables for $\hat{x}'_i=\ReLU(x'_i)$ as distinct variables, and relax only the latter.

	This approach has a similar disadvantage as the model in the previous section: the output computed by the network on the optimized inputs may not equal the output value in the solution. Similarly, it is meaningful to compute the upper bound using this approach.

	\subsection{A Simplified model}

	To avoid above problem, we introduce a simplified model that is to use one variable $y_i$ to represent the difference of each two variables: that is, if $x_i$ and $x'_i$ are two variables in $\mathcal{M}$ and $\mathcal{M}'$ representing the same node, then we set $y_i=x_i-x'_i$ and and $\hat{y}_i=\hat{x}_i-\hat{x}_i'$. The relation between $y_i$, $x'_i$ and $\hat{y}_i$ is $\hat{y}_i = \ReLU(x'_i+y_i)-\ReLU(x'_i).$

	Given  $\gamma_i$ be the upper bound of $y_i$, the constraints for $\hat{y}_i$ are the follows:\begin{align*}
		\hat{y}_i &\leq a \gamma_i               &\quad \hat{y}_i &\geq y_i - a \gamma_i \\
		\hat{y}_i &\geq (a-1) \gamma_i           &\quad \hat{y}_i &\leq y_i + (1-a) \gamma_i
	\end{align*} where $a$ is a binary variable.

	The following plot illustrates the constraints described above.

	\begin{figure}[b!]
		\centering
	\hspace*{10ex}\begin{tikzpicture}
		\begin{axis}[
			xlabel={$y_i$},
			ylabel={$\hat{y}_i$},
			xmin=-2, xmax=2,
			ymin=-2, ymax=2,
			axis lines=center,
			samples=100, 
			unit vector ratio=1 1 1, scale=1, xtick   = {-2,2},
			xticklabels = {$-\gamma_i$,$\gamma_i$},
			yticklabels = {},
			]
			\addplot[blue, thick, fill=blue, fill opacity=0.4] {x} \closedcycle; 
			\addplot[blue, thick] {0};

			\addplot[only marks, mark=*, mark size=2pt, blue] coordinates {(-2,-2)};
			\node[label={above:$(-\gamma_i,-\gamma_i)$}] at (axis cs: -1.35, -2.1) {};

			\addplot[only marks, mark=*, mark size=2pt, blue] coordinates {(2,2)};
			\node[label={above:$(\gamma_i,\gamma_i)$}] at (axis cs: 1.4, 1.5) {};
		\end{axis}
	\end{tikzpicture}
\caption{The possible values of $\val(\hat{y}_i)=\ReLU(\val(x_i))-\ReLU(\val(x'_i))$ depending on $\val(y_i) = \val(x_i)-\val(x'_i)$.}
	\label{fig.1v}
\end{figure}

	Based on above constraints, we can sketch this simplified model:
	\begin{enumerate}
		\item For each input node, each output node, and each pre-activation and post-activation node in the hidden layers,  set one variable $y_i$. 
		\item Set constraints for input nodes.
		\item Set linear constraints . In this case, since the meaning of $y_i$ is $x_i-x'_i$, this constraints will not use the bias.
		\item Between pre- and post- activation nodes, set the MILP constraint described above.
	\end{enumerate}

	The key point is that, although this model sets 3 variables (and their binary variables) for each node in the network, only $y_i$  contributes to the final results, and we can ignore $x_i,x_i'$ (and their binary variables) during the optimization.

	As a result, we can relax the binary variables used to $\hat{x}_i = \ReLU(x_i)$ and $\hat{x}'_i = \ReLU(x'_i)$.

	Now the simplified model contains the same number of binary variables as the MILP model for local robustness. Hence this model runs much faster compared to the previous one. The trade-off is lower accuracy compared to the first model under unlimited time. In practice, with a reasonable timeout, this simplified model can usually obtain a better bound.

	One major disadvantage of this model is that the solution obtained through optimization may not be valid-i.e., the output computed by the network on the optimized inputs may not equal the output value in the solution. Nevertheless, it is meaningful to compute the upper bound.

	%	However, in practice, with a reasonable timeout, this simplified model can usually obtain a better bound.
	%