### 2025-07-29 AAAI2026/a_trick.tex

\subsection*{A trick of $L_1$ norm constraints on input}

A trick to set $L_1$ norm constraints on the input space into linear constraints is described below.

Suppose the dimension of input space is $n$, and $x_i,i\leq n$ are the input variables, $\vec{x}$ is the input vector, and we hope to set a constraints of $\vec{x}$ that $\|\vec{x}\| \leq c$,  then we can set the constraints by add variables $A_i,i<n$ with the only constraints \begin{align}\label{L1constraint}\begin{cases}
	A_i \geq x_i &\text{ for }i \leq n\\ 
	A_i\geq -x_i &\text{ for }i \leq n\\
	\Sigma_{i\leq n} A_i \leq c	&\end{cases}
\end{align}

We show that this constraint is equivalent to $\|\vec{x}\| \leq c$.

The first two lines of constraints \ref{L1constraint} is equivalent to that $A_i\geq |x_i|$. So the whole constraints is non-weaker than $\|\vec{x}\| \leq c$.

Next, for every instance that satisfy $\|\vec{x}\| \leq c$, then there is also an instance which satisfy constraints \ref{L1constraint}: simply let all $A_i = |x_i|$. And this shows that $\|\vec{x}\| \leq c$ is non-weaker than constraints \ref{L1constraint}. Therefore they are equivalent.

\subsection*{A proof of constraints}

We will show that the constraints corresponding to \ref{fig.2v} is exact.

We only need to consider all cases of binary variable. First, we know the constraints for $\hat{x_i}'=\ReLU(x_i')$ are exact.

We have 2 binary variables and 4 cases in total. We only need to check that, in all 4 cases, $$\hat{y}_i = \ReLU(x'_i+y_i)-\ReLU(x'_i).$$

Case 1: if $a = 1$ and $a' = 1$, then $x'_i \geq 0 $ and $y_i+x'_i\geq 0$, then we need to show $\hat{y}_i = y_i$ based on  $\hat{x'_i} = x'_i$. This is true by the two inequalities in line 4.

Case 2: if $a = 1$ and $a' = 0$, then $x'_i \leq 0 $ and $y_i+x'_i\geq 0$, then we need to show $\hat{y}_i = y_i+x'_i$ based on  $\hat{x'_i} = 0$. This is true by the two inequalities in line 6.

Case 3: if $a = 0$ and $a' = 0$, then $x'_i \leq 0 $ and $y_i+x'_i\leq 0$, then we need to show $\hat{y}_i = 0$ based on  $\hat{x'_i} = 0$. This is true by the two inequalities in line 3.

Case 4: if $a = 0$ and $a' = 1$, then $x'_i \geq 0 $ and $y_i+x'_i\leq 0$, then we need to show $\hat{y}_i = -x'_i$ based on  $\hat{x'_i} = x'_i$. This is true by the two inequalities in line 5.