### 2025-08-02 AAAI2026/related.tex
full paper

\section{Related Work}

Some works \cite{Leino,Zhang22a,Sun22,Chen21,REGLO} focus on training a network to be - rather than proving the DNN is - globally robust.
Other works \cite{Bastani16,Ruan19,Gopinath18} estimate global robustness bounds, but without hard verification certificates, e.g. using probabilistic guarantees 
\cite{Levy23,Mangal19}.

Closer to our paper, several works  \cite{Marabou,lipshitz,GROCET} consider a global certified restricted Lipschitz bound, to understand how the output change with input perturbations. However, they do not consider whether the classification changes.

The most related work is Vhagar \cite{vhagar}, which also computes $\beta^\varepsilon_{i,j}$ bounds, but limited to $L_\infty$ perturbations (and variants). They do not however report real-time certified robust percentage, as in Table \ref{table.cert}. Last, their hyper-adversarial attack to find better lower bounds seems very efficient, and we will consider it in the future.

Technically, none of the MILP base work \cite{vhagar,ITNE} 
consider $L_1$ bounds. Further, they use the classical MILP encoding from \cite{MILP}, and not our novel "2v" model, nor even our "1v" model. However, \cite{ITNE} 
defines the {\em diff variables} $y,\hat{y}$, adding explicitely the linear relaxation $\frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}$ implied by our models, Eq. (3) in \cite{ITNE} corresponding essentially to it. \cite{vhagar} does things differently, adding per neuron constraints (depending on the perturbation) between the (binary) variables to simplify the MILP problem, which is well-adapted to occlusion and patches but less to $L_1$ perturbations.