### 2025-08-05 AAAI2026/PCA.tex
images as PDF

\section{Explanations on PCA model order reduction}

Some words on PCA, learn of 2 Matrix full-dim to reduced-dim and back.

Then how we use it (pictures and text).

\begin{figure*}[]
	\centering
	\includegraphics[scale=0.9]{MNIST.pdf} \hspace{1.5cm}
	\caption{MNIST.}
	\label{fig.MNIST}
\end{figure*}

\begin{figure*}[]
	\centering
	\includegraphics[scale=0.9]{PIPE.pdf} \hspace{1.5cm}
	\caption{PIPE.}
	\label{fig.PIPE}
\end{figure*}