### 2025-11-10 CAV 2026/CAV2026_main.tex
Add CAV2026 folder

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}

\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet}  % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
%
% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
\usepackage{algorithm}
\usepackage{algorithmic}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
\usepackage{lineno}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{cases}
\captionsetup{compatibility=false}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
%\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}
\usepackage{newfloat}
\usepackage{listings}
\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%\newtheorem{proposition}{Proposition}
%\newtheorem{definition}{Definition}
\newcommand{\vW}{\boldsymbol{W}}
\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}
\newcommand{\Improve}{\mathrm{Improve}}
\newcommand{\Utility}{\mathrm{SAS}}
\newcommand{\Sol}{\mathrm{Sol}}
\newcommand{\sol}{\mathrm{sol}}
\newcommand{\UB}{\mathrm{UB}}
\newcommand{\LB}{\mathrm{LB}}
\newcommand{\ub}{\mathrm{ub}}
\newcommand{\lb}{\mathrm{lb}}
\newcommand{\B}{\mathrm{B}}
\usepackage{amsmath, amssymb, amsfonts}
\newcommand{\ReLU}{\mathrm{ReLU}}
\newcommand{\CMP}{{\textrm{CMP}}\ }
\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\toolname}{Hybrid MILP}

% REMOVE THIS: bibentry
% This is only needed to show inline citations in the guidelines document. You should not need it and can safely delete it.
\usepackage{bibentry}
% END REMOVE bibentry

\begin{document}

\title{Order reduction and MILP models for real-time DNN robustness certification.}
\date{}

\maketitle

\begin{abstract}
	\input{abstract.tex}
	\iffalse
	Most DNNs are brittle to small perturbations. Extensive works have thus been performed to verify robustness for DNNs.
	However, these works mostly consider local robustness, i.e. in the neighborhood of an image.
	While local robustness is useful to have an idea how often non robust images happen, by repeating the verification on 1000 or 10000 pre-obtained images, the main shortcoming is that we have no guarantee that a specific new incoming image, e.g. in a video feed, is robust: The verification process takes too long and requires too much resources to be performed online on embedded systems.

	In this paper, we consider {\em global} robustness, that is, guarantees not restricted to a set of local images. For that, we consider {\em bounds} on the switch of values between the different decision classes of a DNN due to a given perturbation. 
	The verification question is much harder than local robustness, as the number of complex variables doubles (from the deviation image to the image and its deviation).
	Further, the values each neuron can take is no more in a small neighborhood.
	Therefore, the global verification process is very complex.
	To obtain useable bounds, we develop several novel partial MILP models for global robustness, with different trade-offs. Last, we use order reduction techniques to reduce the space of images considered, avoiding unrealistic inputs, by using linear PCA. 
	This results into usable bounds, allowing in real time to certify robustness for $87\%$ of incoming images in the MNIST benchmark for a L1-perturbation of $0.5$, as well as for a surrogate computing the hidden plastic strain associated to a deformation map of a pipe.
	\fi
\end{abstract}

\section{Introduction}

\input{introduction}

\input{notation}

\input{global}

\input{a_trick}

\newpage

\input{experiments}

%\vspace{-0.5cm}
\input{related}

\section{Conclusion}

In this paper, we introduced a novel "2v" MILP model for global verification questions, which is 4 to 8 times more accurate than the classical MILP model,
and more accurate than the ITNE model of \cite{lipshitz,ITNE}, see 
Table \ref{table.classical}. Using our "3v" or "1v" models abstracting away some features of the "2v" model,
even better bounds are reached (Tables 3, 4, 5). We also explained how to support $L_1$-perturbations efficiently on top of classical $L_\infty$-perturbations (Prop \ref{prop.l1}). To further lower the bounds by $> 10$ times, we resort to model order reduction, reducing the space of search closer to the space of the training dataset, without sacrificing DNN accuracy ($97\%$). This results into, to the best of our knowledge, the first {\em real-time} robustness certification reported for a DNN, with $86\%$ of robustness for an $L_1$-perturbation of $.5$ and $53\%$ of robustness for an $L_1$-perturbation twice as large of $1$. Still, global verification is only at its beginning, and combining different methods is necessary to scale to more complex DNNs.

\newpage

\bibliographystyle{plain} 
\bibliography{references}
\end{document}